### 2021-07-08 latex_files/ch1/1_4_online_notebooks.tex

Since the dissertation is mainly focused on the implementation of the
ROMC, we have created some interactive notebooks\footnote{We have used
  the jupyter notebooks \autocite{Kluyver:2016aa}} supporting the main
document. The reader is advised to exploit the notebooks in order to
(a) review the code used for performing the inference (b) understand
the functionalities of our implementation in a practical sense (c)
check the validity of our claims and (c) interactively execute
experiments.

The notebooks are provided in two formats:\footnote{The
  notebooks in the two repositories are identical.}

\begin{itemize}
\item The first format is located in
  \href{https://github.com/givasile/edinburgh-thesis/tree/master/notebook_examples}{this}
  Github repository. In order for the reader to experiment with the
  examples, they should clone the repository into their pc, execute
  the installation instructions and run the notebooks using the
  Jupyter package.
\item The second format is the google colab notebook
  \autocite{Bisong2019}. Following the links below, the reader can
  view the notebook and make an online clone of it. Hence, without any
  installation overhead, they may interactivelly explore all the
  provided functionalities.
\end{itemize}

The notebooks provide a practical overview of the implemented
functionalities and it are easy to use; in particular, the google
colab version is entirely plug-and-play. Therefore, we encourage the
reader of the dissertation to use them as supporting material.  The
following list contains the implemented examples along with their
links:

\begin{itemize}
\item Simple 1D example
  \begin{itemize}
  \item \href{https://colab.research.google.com/drive/1lGRp0XrNfZ64NN0ASB_tYEKowXwlveDC}{Google colab}
  \item \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_1D.ipynb}{Github repository}
  \end{itemize}

\item Simple 2D example
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1Fof_WmCi1YizzSI_63aEsbLXsno5gSZ3}{Google colab}
  \item \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_2D.ipynb}{Github repository}
  \end{itemize}

\item Moving Average example  
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1nkdACQ370SSc0KB1bHv4sBRaxMlMqoNH}{Google colab}
  \item
    \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_MA2.ipynb}{Github repository}
  \end{itemize}

\item Extensibility example
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v}{Google colab}
  \item
    \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/extending_romc.ipynb}{Github repository}
  \end{itemize}

\end{itemize}