### 2021-07-08 latex_files/ch2/2_2_ROMC.tex

The simplifications introduced by OMC, although quite useful from a
computational point-of-view, they suffer from some significant failure modes:

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$, for each nuisance variable,
  shrinks to a single point $\thetab_i^*$; this simplification may add
  significant error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. This approach is error prone at many cases
  e.g.\ when $g_i$ is almost flat at $\thetab_i^*$, leading to a
  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0 \Rightarrow w_i
  \rightarrow \infty$, thus dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

\subsubsection{Sampling and computing expectation in ROMC}

The ROMC approach resolves the aforementioned issues. Instead of
collapsing the acceptance regions $\mathcal{S}_i$ into single points,
it tries to approximate them with a bounding box.\footnote{The
  description on how to estimate the bounding box is provided in the
  following chapters.}. A uniform distribution is then defined on the
bounding box area, used as the proposal distribution for importance
sampling. If we define as $q_i$, the uniform distribution defined on
the $i-th$ bounding box, weighted sampling is performed as:

\begin{gather}
  \label{eq:sampling}
  \thetab_{ij} \sim q_i \\
  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
\end{gather}

\noindent
Having defined the procedure for obtaining weighted samples, any
expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
as,

\begin{equation} \label{eq:expectation}
  E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}

\subsubsection{Construction of the proposal region}

In this section we will describe mathematically the steps needed for
computing the proposal distributions $q_i$. There will be also
presented a Bayesian optimisation alternative when gradients are not
available.

\subsubsection*{Define and solve deterministic optimisation problems}

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$ a
deterministic function is defined as
$f_i(\thetab) = M_d(\thetab,\vb_i)$. For constructing the proposal
region, we search for a point
$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$; this point can be
obtained by solving the the following optimisation problem:

\begin{subequations}
\begin{alignat}{2}
&\!\min_{\thetab}        &\qquad& g_i(\thetab) = d(\data,  f_i(\thetab))\label{eq:optProb}\\
&\text{subject to} &      & g_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
We maintain a list of the solutions $\thetab_i^*$ of the optimisation
problems. If for a specific set of nuisance variables $\vb_i$, there
is no feasible solution we add nothing to the list. The optimisation
problem can be treated as unconstrained, accepting the optimal point
$\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$ only if
$g_i(\thetab_i^*) < \epsilon$.

\subsubsection*{Gradient-based approach}
\label{subsubsec:GB_approach}

The nature of the generative model $M_r(\thetab)$, specifies the
properties of the objective function $g_i$. If $g_i$ is continuous
with smooth gradients $\nabla_{\thetab} g_i$ any gradient-based
iterative algorithm can be used for solving~\ref{eq:optProb}. The
gradients $\nabla_{\thetab} g_i$ can be either provided in closed form
or approximated by finite differences.

\subsubsection*{Bayesian optimisation approach}
\label{subsubsec:GP_approach}

In cases where the gradients are not available, the Bayesian
optimisation scheme provides an alternative
choice~\cite{Shahriari2016}. With this approach, apart from
obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
the distance $g_i$ is fitted; this approximate model can be used in
the following steps for making the method more
efficient. Specifically, in the construction of the proposal region
and in
equations~\eqref{eq:approx_posterior},~\eqref{eq:sampling},~\eqref{eq:expectation}
it could replace $g_i$ in the evaluation of the indicator
function, providing a major speed-up.

\subsubsection*{Construction of the proposal area $q_i$}

After obtaining a $\thetab_i^*$ such that
$g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
around it. The bounding box $\mathcal{\hat{S}}_i$ must contain the
acceptance region around $\thetab_i^*$, i.e.\
$\{ \thetab : g_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The second condition
$d(\thetab, \thetab_i^*) < M$ is meant to describe that if
$\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
number of disjoint sets of $\thetab$ that respect
$g_i(\thetab) < \epsilon$, we want our bounding box to fit only the
one that contains $\thetab_i^*$. We seek for a bounding box that is as
tight as possible to the local acceptance region (enlarging the
bounding box without a reason decreases the acceptance rate) but large
enough for not discarding accepted areas.

In contrast with the OMC approach, we construct the bounding box by
obtaining search directions and querying the indicator function as we move on them. The
search directions $\mathbf{v}_d$ are computed as the eigenvectors of
the curvature at $\thetab_i^*$ and a line-search method is used to
obtain the limit point where
$g_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite direction along the search line}. The Algorithm~\ref{alg:region_construction} describes the method in-depth. After the limits are obtained along all
search directions, we define bounding box and the uniform distribution $q_i$. This is the proposal distribution used for the importance
sampling as explained in \eqref{eq:sampling}.

\subsubsection*{Fitting a local surrogate model $\hat{g}_i$}

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in the surface outside the box. In the future
steps (e.g. sampling, evaluating the posterior) we will only evaluate
$g_i$ inside the corresponding bounding box. Hence, we could fit a
local surrogate model $\hat{g}_i$ for representing the local area
around $\theta_i^*$. Doing so, in the future steps we can exploit
$\hat{g}_i$ for evaluating the indicator function instead of running
the whole deterministic simulator.

Any ML regression model may be chosen as local surrogates. The choice
should consider the properties of the local region (i.e. size,
smoothness). The ROMC proposes fitting a simple quadratic model. The
training set $X: \mathbb{R}^{N \times D}$ is created by sampling $N$
points from the bounding box and the labels $Y: \mathbb{R}^{N}$ are
the computed by evaluating $g_i$. The quadratic model is fitted on the
data points, for minimising the square error.

This additional step places an
additional step in the training part, increasing the computational
demands, but promises a major speed at the inference phase (sampling,
posterior evaluation). It is frequent in ML problems, to be quite
generous with the execution time at the training phase, but quite
eager at the inference phase. Fitting a local surrogate model aligns
with this requirement.