### 2021-07-08 latex_files/ch3/3_1_1_general_design.tex

In figure \ref{fig:romc_overview} we present an overview of our
implementation; one may interpret figure~\ref{fig:romc_overview} as a
depiction of the main class of our implementation, called
\pinline{ROMC}, while the entities inside the green and blue
ellipses are the main functions of the class. Following Python's
naming principles, the methods starting with an underscore (green
ellipses) represent internal (private) functions and are not meant to
be used by a user, whereas the rest of the methods (blue ellipses) are
the functionalities the user interacts with. As mentioned before, the
implementation favours extensibility; the building blocks that compose
the method have been designed in a modular fashion so that a
practitioner may replace them without the method to collapse.

Figure \ref{fig:romc_overview} groups the ROMC implementation into the
training, the inference and the evaluation part. The training part includes all the
steps until the computation of the proposal regions; sampling the
nuisance variables, defining the optimisation problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalised posterior (and
the normalised one, in low-dimensional cases), sampling and computing
an expectation. Moreover, the ROMC implementation provides some
utilities for inspecting the training process, such as plotting the
histogram of the distances
$d^*_i = g_i(\theta_i^*), \: \forall i \in \{1, \ldots, n_1 \}$ after
solving the optimisation problems and visualising the constructed
bounding box\footnote{if the parametric space is up to $2D$}. Finally,
two functionalities for evaluating the inference are implemented; (a)
computing the Effective Sample Size (ESS) of the weighted samples and
(b) measuring the divergence between the approximate posterior the
ground-truth, if the latter is available.\footnote{Normally, the
  ground-truth posterior is not available; However, this functionality
  is useful in cases where the posterior can be computed numerically
  or with an alternative method (i.e.\ ABC Rejection Sampling) and we
  would like to measure the discrepancy between the two
  approximations.}

\subsubsection*{Parallelising the processes}

As stated above, the most critical advantage of the ROMC method is that it can be fully parallelised. In our implementation, we have exploited this fact by implementing a parallel version in the following tasks; (a) solving the optimisation problems, (b) constructing bounding box regions, (c) sampling and (d) evaluating the posterior. Parallelism has been achieved using the package \pinline{multiprocessing}. The specific package enables concurrency, using subprocesses instead of threads, hence side-stepping the Global Interpreter (GIL). In our implementation we use the \pinline{Pool} object, which offers a convenient means of parallelising the execution of a function across multiple input values (data parallelism). For activating the parallel version of the algorithms, the user has to just pass the argument \pinline{parallelize=True} at the initialisation of the \pinline{ROMC} method.

\subsubsection*{Simple one-dimensional example}

For illustrating the functionalities we choose as running example the
following model, introduced by~\autocite{Ikonomov2019},

\begin{gather} \label{eq:1D_example} p(\theta) =
\mathcal{U}(\theta;-2.5,2.5)\\ p(y|\theta) = \left\{
    \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
\\ |\theta| - c + u & \mbox{otherwise}
    \end{array} \right.\\ u \sim \mathcal{N}(0,1)
\end{gather}

\noindent

In the model \eqref{eq:1D_example}, the prior is the uniform
distribution in the range $[-2.5, 2.5]$ and the likelihood a Gaussian
distribution. There is only one observation $y_0 = 0$. The inference
in this particular example can be performed quite easily, without
incorporating a likelihood-free inference approach. We can exploit this
fact for validating the accuracy of our implementation. The
ground-truth posterior, approximated computationally, is shown in
figure \ref{fig:example_gt}.

\begin{figure}[h]
    \begin{center}
\includegraphics[width=0.7\textwidth]{./Thesis/images/chapter3/example_gt.png}
    \end{center}
  \caption[Ground-truth posterior distribution of the simple 1D
example.]{Ground-truth posterior distribution of the simple 1D
example.}
  \label{fig:example_gt}
\end{figure}

\subsubsection*{ELFI code for modelling the example}

In the following code snippet, we code the model at \textit{ELFI} and
we initialise the ROMC inference method. We observe that the
initialisation of the ROMC inference method is quite intuitive; we
just pass the final (distance) node of the simulator as argument, as
in all $\textit{ELFI}$ inference methods. The argument
\pinline{bounds}, although optional, is important for many
functionalities (e.g. approximating the partition function, setting
the bounds of the Bayesian optimisation etc.) so it is recommended to
be passed.

\begin{pythoncode}
  import elfi import scipy.stats as ss
  import numpy as np

  def simulator(t1, batch_size=1,random_state=None):
      if t1 < -0.5:
          y = ss.norm(loc=-t1-c, scale=1).rvs(random_state=random_state)
      elif t1 <= 0.5:
          y = ss.norm(loc=t1**4, scale=1).rvs(random_state=random_state)
      else:
          y = ss.norm(loc=t1-c, scale=1).rvs(random_state=random_state)
      return y

  # observation
  y = 0

  # Elfi graph
  t1 = elfi.Prior('uniform', -2.5, 5)
  sim = elfi.Simulator(simulator, t1, observed=y)
  d = elfi.Distance('euclidean', sim)

  # Initialise the ROMC inference method
  bounds = [(-2.5, 2.5)] # limits of the prior
  parallelize = True # activate parallel execution
  romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
\end{pythoncode}

\begin{figure}[!ht]
    \begin{center}
\includegraphics[width=0.8\textwidth]{./Thesis/graphs/ROMC.png}
    \end{center}
    \caption[Overview of the ROMC implementation.]{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
    \label{fig:romc_overview}
\end{figure}