### 2021-07-08 latex_files/ch3/3_1_user.tex

In section \ref{subsec:general_design} we present an overview of the implementation. Afterwards, we divide the functionalities in 3 parts; in section \ref{subsec:training} we present the methods used for training (fitting) the model, in section \ref{subsec:inference} those used for performing the inference and in section \ref{subsec:evaluation} those used for evaluating the approximate posterior and the obtained samples.

\subsubsection{General design} 
\label{subsec:general_design}
\input{Thesis/ch3/3_1_1_general_design}

\subsubsection{Training part} 
\label{subsec:training}
\input{Thesis/ch3/3_1_2_training}

\subsubsection{Inference part} 
\label{subsec:inference}
\input{Thesis/ch3/3_1_3_inference}

\subsubsection{Evaluation part} 
\label{subsec:evaluation}
\input{Thesis/ch3/3_1_4_evaluation}