### 2021-07-17 latex_files/ch3/3_1_ELFI.tex

\pkg{Engine for Likelihood-Free Inference (ELFI)}\footnote{Extended
  documentation can be found
  \href{https://elfi.readthedocs.io/en/latest/}{here}}~\cite{1708.00707}
is a \proglang{Python} package dedicated to Likelihood-Free Inference
(LFI). We implemented ROMC at \pkg{ELFI} since it provides convenient
modules for all the fundamental components of a probabilistic model
(e.g. prior, simulator, summaries etc.). Furthermore, \pkg{ELFI}
already supports some recently proposed likelihood-free inference
methods, making it straightforward to perform comparisons between
methods. The implementation was officially added to \pkg{ELFI} at
version \pkg{v0.8.0}.

% \subsubsection{Modelling at ELFI}
% \label{sec:modelling}

\pkg{ELFI} models the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package \pkg{NetworkX},
which supports general-purpose graphs. In most cases, the structure of
a likelihood-free model follows the pattern of figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential sampling. \pkg{ELFI} automatically considers as parameters
of interest, i.e. those we try to infer a posterior distribution,
as the ones  defined through the \code{elfi.Prior} class. Finally, the
observations are passed to the appropriate node through the argument
\code{observed}.

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
    \end{center}
    \caption[Baseline example for creating an \pkg{ELFI} model]{Baseline example for creating an \pkg{ELFI} model. Image taken from \cite{1708.00707}}
    \label{fig:elfi}
\end{figure}

% \subsubsection{Inference at ELFI}
% \label{sec:inference-methods}

All inference methods at \pkg{ELFI} share the following rules;

\begin{itemize}
\item they follow the signature \code{elfi.<Class name>(<output node>,
    *arg)}; the initial argument is the output node model and the rest
  of the arguments are hyper-parameters of the method.
\item they provide a central sampling functionality, which in most
  cases is named \\ \code{<method_name>.sample()}.
\end{itemize}