### 2021-07-23 latex_files/ch2/2_2_LFI_background.tex
create tikz figures and plots

In simulator-based models we cannot evaluate the posterior
$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood,

\begin{equation} \label{eq:likelihood}
  L(\thetab) =
  \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in B_{d,\epsilon}(\data))
\end{equation}
%
where $c_\epsilon$ is a proportionality factor dependend on
$\epsilon$ and $B_{d,\epsilon}(\data)$ is a region around $\data$.

Intuitively, equation~\ref{eq:likelihood} describes that the
likelihood of a specific parameter configuration $\thetab$ is equal to
the probability that the simulator will produce outputs infinitively
close to the observations $\data$, using this configuration.

For continuous cases, the probability of replicating the data
infinitely close becomes zero. In this scenario a relaxation must be
introduced by accepting simulated data that fall close to the data
i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
$\epsilon > 0$. The region can be
defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
where $d(\cdot, \cdot)$ can represent any valid distance.

Such relaxation, introduces the approximate likelihood,

\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = Pr(\yb \in \region(\data)), \text{where  } \epsilon > 0
\end{equation}

and consequently the approximate posterior,

\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab)
\end{equation}