### 2021-07-23 latex_files/images/chapter3/example_marginal.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
tick align=outside,
tick pos=left,
title={Histogram of the samples drawn},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta\)},
xmin=-3.3, xmax=3.3,
xtick style={color=black},
xtick={-4,-3,-2,-1,0,1,2,3,4},
xticklabels={
  \(\displaystyle {\ensuremath{-}4}\),
  \(\displaystyle {\ensuremath{-}3}\),
  \(\displaystyle {\ensuremath{-}2}\),
  \(\displaystyle {\ensuremath{-}1}\),
  \(\displaystyle {0}\),
  \(\displaystyle {1}\),
  \(\displaystyle {2}\),
  \(\displaystyle {3}\),
  \(\displaystyle {4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=0.6,
ytick style={color=black},
ytick={0,0.1,0.2,0.3,0.4,0.5,0.6},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.1}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.3}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {0.6}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-3,0) rectangle (axis cs:-2.91428571428571,0);
\draw[draw=none,fill=color0] (axis cs:-2.91428571428571,0) rectangle (axis cs:-2.82857142857143,0);
\draw[draw=none,fill=color0] (axis cs:-2.82857142857143,0) rectangle (axis cs:-2.74285714285714,0);
\draw[draw=none,fill=color0] (axis cs:-2.74285714285714,0) rectangle (axis cs:-2.65714285714286,0);
\draw[draw=none,fill=color0] (axis cs:-2.65714285714286,0) rectangle (axis cs:-2.57142857142857,0);
\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.00722007830101246);
\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0252702740535435);
\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0347466268236223);
\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0374541561865022);
\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.0510691609826969);
\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0647744167576544);
\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.080503873056289);
\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.089941546835469);
\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.113625982262183);
\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.12614508231626);
\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.145794009692587);
\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.171554217630842);
\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.157720031886223);
\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.174635643905738);
\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.199983275941078);
\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.216744171997);
\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.240854076323595);
\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.277844084619317);
\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.253682608304859);
\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.2729834247631);
\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303178823657691);
\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.297596155971373);
\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.292864426084818);
\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.327185584008558);
\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.303294860630386);
\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.313867118142585);
\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.316600433499394);
\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.325419243424203);
\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.322737500055257);
\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.331556309980064);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.327533694926642);
\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.351940138183459);
\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.308245771465366);
\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.329080854562574);
\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.311752666640144);
\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.312139456549128);
\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.322673035070425);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.302650210782081);
\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.313763974166856);
\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.278656343428181);
\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.269785961515508);
\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.245830773152508);
\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.227754791406045);
\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.218858623499437);
\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.218433154599559);
\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.21010427855946);
\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.193691493421625);
\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.194078283330608);
\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.130644738257426);
\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.150177628661059);
\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.0986572127845487);
\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0960012554095325);
\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0732709017583105);
\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0897223658870459);
\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.0617058834797238);
\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.0434622927727019);
\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0347466268236225);
\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0374541561865018);
\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0320390974607428);
\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00496380383194607);
\draw[draw=none,fill=color0] (axis cs:2.57142857142857,0) rectangle (axis cs:2.65714285714286,0);
\draw[draw=none,fill=color0] (axis cs:2.65714285714286,0) rectangle (axis cs:2.74285714285714,0);
\draw[draw=none,fill=color0] (axis cs:2.74285714285714,0) rectangle (axis cs:2.82857142857143,0);
\draw[draw=none,fill=color0] (axis cs:2.82857142857143,0) rectangle (axis cs:2.91428571428571,0);
\draw[draw=none,fill=color0] (axis cs:2.91428571428571,0) rectangle (axis cs:3,0);
\addplot [semithick, red, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.59322023391724 0
-2.49152541160583 0.0369433164596558
-2.38983058929443 0.0452905893325806
-2.28813552856445 0.0549525022506714
-2.18644070625305 0.0659897327423096
-2.08474564552307 0.0784285068511963
-1.98305082321167 0.0922528505325317
-1.88135588169098 0.107397437095642
-1.77966105937958 0.123741984367371
-1.67796611785889 0.141106963157654
-1.5762711763382 0.159253358840942
-1.27118647098541 0.21515953540802
-1.16949152946472 0.232988834381104
-1.06779658794403 0.249699831008911
-0.96610164642334 0.264856100082397
-0.864406824111938 0.278041839599609
-0.762711882591248 0.288881063461304
-0.661016941070557 0.297054767608643
-0.559322118759155 0.302316904067993
-0.457627058029175 0.304275751113892
-0.355932235717773 0.304529428482056
0.254237294197083 0.304565906524658
0.355932235717773 0.304529428482056
0.457627058029175 0.304275751113892
0.559322118759155 0.302316904067993
0.661016941070557 0.297054767608643
0.762711882591248 0.288881063461304
0.864406824111938 0.278041839599609
0.96610164642334 0.264856100082397
1.06779658794403 0.249699831008911
1.16949152946472 0.232988834381104
1.27118647098541 0.21515953540802
1.3728814125061 0.196650266647339
1.5762711763382 0.159253358840942
1.67796611785889 0.141106963157654
1.77966105937958 0.123741984367371
1.88135588169098 0.107397437095642
1.98305082321167 0.0922528505325317
2.08474564552307 0.0784285068511963
2.18644070625305 0.0659897327423096
2.28813552856445 0.0549525022506714
2.38983058929443 0.0452905893325806
2.49152541160583 0.0369433164596558
2.59322023391724 0
3 0
};
\end{axis}

\end{tikzpicture}