### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_rejection.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Rejection ABC - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.02965548855742);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.17070418562009);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.6643746253394);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.50922105857047);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.85479036637399);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.25677915300258);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.3484608060933);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.53887654712789);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.74339715786876);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.79981663669382);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.02549455199407);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.87739342007828);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.67287280933742);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.11012377023167);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.97612750802216);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.77865933213442);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.67992524419055);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.68697767904369);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.24972671814944);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,2.45424732889032);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,2.35551324094642);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,2.12278289079306);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.9464720194647);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.960576889171);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.5162734934236);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.25533340385768);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.42459184033288);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.10723227194188);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.853344617229104);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.768715398991495);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.66292887619451);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.550089918544377);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.500722874572446);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.430198526041119);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.352621742656649);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.373779047216054);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.282097394125319);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.148101131915795);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.112838957650128);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.0916816530907302);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.112838957650129);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0634719136781969);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0493670439719316);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0423146091187979);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0564194788250647);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.00705243485313299);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0.0141048697062662);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0.0282097394125324);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0.014104869706266);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0.00705243485313299);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.51592730438119 0
0.51592730438119 3.5
};
\addlegendentry{$\mu = 0.516$}
\addplot [semithick, black, dashed]
table {%
0.336596813861936 1
0.648443255776683 1
};
\addlegendentry{$\sigma = 0.142$}
\end{axis}

\end{tikzpicture}