### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_romc.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.02018167274615);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.39039538321726);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.64837235793467);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.83766715107148);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.03114988210955);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.20118016090059);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.42565363240795);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.83858430947187);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.02313277298727);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.49050664276752);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.58012851385441);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.63066296452957);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.34895434170502);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.09544450074894);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.92569341781802);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.63923846537203);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.52616414203809);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.21569834562492);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.07749639488345);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.95995493778819);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.85302292337607);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.59560434037885);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.66428652195943);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.3535415296862);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.45042249313527);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.1846280343355);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.19830863147962);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.979977468894081);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.941727636062617);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.705807134291806);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.614789283904501);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.496689435089055);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.456206035376906);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.425773686627776);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.449226138874808);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.309907404693004);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.224752667367446);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.192924339317897);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.0974393551692405);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.176730979433038);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.170309474651111);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.145461043103651);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0304323487491329);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0206604936462);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0217772770865351);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.503688298846839 0
0.503688298846839 3.5
};
\addlegendentry{$\mu = 0.504$}
\addplot [semithick, black, dashed]
table {%
0.323676768279387 1
0.634437489183658 1
};
\addlegendentry{$\sigma = 0.141$}
\end{axis}

\end{tikzpicture}