### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.67339010906917);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,2.12285439456296);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.41994534810638);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.13421448089961);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.60269456308737);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.61479552461991);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.51971654115006);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.94571977877467);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.78840727885186);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.70814579929938);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.89534026545558);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.97807132899427);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.47304140381032);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.18508791101595);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.50193553644923);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.27646251850645);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,1.84996536408459);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.10951864103732);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.94010517958197);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.70154336651221);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.88058820551382);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.72549833117864);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.29579071757466);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.50076618843174);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.46940247180661);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.16860714228384);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.37012519556019);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.767793661318775);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.885098900664694);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.65715629873569);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.82731063538692);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.813727923462651);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.801873920328749);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.459095663040085);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.672961636247559);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.561583398468614);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.382538559466961);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.340555631701063);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.294127452759942);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.239055729867026);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.283755200017783);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.176575255015415);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.283261283220537);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.258071526560991);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0775449371676099);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0155583791132463);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0.0172870879036073);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.500240536632723 0
0.500240536632723 3.5
};
\addlegendentry{$\mu = 0.500$}
\addplot [semithick, black, dashed]
table {%
0.293929386789429 1
0.656599793802562 1
};
\addlegendentry{$\sigma = 0.165$}
\end{axis}

\end{tikzpicture}