### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_rejection.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Rejection ABC - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0.016001600160016);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.02000200020002);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0600060006000599);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0800080008000799);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.132013201320132);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.216021602160216);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.344034403440344);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.412041204120412);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.508050805080508);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.832083208320833);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.884088408840883);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.21612161216122);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.39213921392139);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.71217121712171);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.95219521952195);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.07620762076207);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.15221522152216);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.41624162416241);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.49224922492249);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.48824882488249);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.30823082308231);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.30423042304231);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.2042204220422);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.90819081908191);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.64816481648165);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.50815081508151);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.3001300130013);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.964096409640963);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.920092009200919);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.744074407440743);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.556055605560556);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.472047204720474);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.38003800380038);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.228022802280228);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.28002800280028);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.192019201920192);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.140014001400141);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0960096009600955);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0880088008800883);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0560056005600562);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0760076007600756);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0280028002800281);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0360036003600358);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.0240024002400241);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0280028002800279);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0.00800080008000803);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0.0200020002000201);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0.00400040004000398);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0.0280028002800281);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0.00400040004000398);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0.0160016001600161);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0.0120012001200119);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0.00800080008000796);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0.00800080008000803);
\addplot [semithick, red, dashed]
table {%
0.0698328247281109 0
0.0698328247281109 3
};
\addlegendentry{$\mu = 0.070$}
\addplot [semithick, black, dashed]
table {%
-0.136927686891309 1
0.240559901293153 1
};
\addlegendentry{$\sigma = 0.172$}
\end{axis}

\end{tikzpicture}