### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_romc.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0132189786765283);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0698717444330784);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.203477850342276);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.355023998741048);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.514595955621997);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.712880635769923);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.916830592493502);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.952710677472651);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.1953733574632);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.3804390589346);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.52584782437642);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.59477535604688);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.7604846958848);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.91108663152097);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.09048705641672);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.27366433236291);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.32654024706902);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32811393500671);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.26296325438667);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.10968604925645);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15437878668663);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.78361790856873);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.58785112912109);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44590447714217);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.22086710205366);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.01125186875443);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.805413486505622);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.800220316311271);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.62758674954756);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.600834054606967);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.396254622708316);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.300574396097251);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.197497836179084);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.168384609331968);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.156581949799354);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.068455425289165);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0550790778188679);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.066094893382642);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0275395389094342);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0275395389094339);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.0308218751896741 0
0.0308218751896741 3
};
\addlegendentry{$\mu = 0.031$}
\addplot [semithick, black, dashed]
table {%
-0.179871006878183 1
0.197679132295466 1
};
\addlegendentry{$\sigma = 0.172$}
\end{axis}

\end{tikzpicture}