### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0710656818158832);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0516473550957937);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.194048417709784);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.2505503544856);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.348585934523831);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.648086654282988);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.583763447022691);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.929922090706508);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.79776958941701);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,0.987907371884556);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.05978215064711);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.2930717702704);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.55481463251829);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.01101046094927);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.35069632905971);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.34206596162857);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.80716185647295);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.71681269742808);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.65275918915001);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.45763197551135);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.24848041479702);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.8837125413398);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.80455589005721);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.47147764701123);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.35887832193294);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.990739211197902);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.834853199472739);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.568255755544842);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.531441844471339);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.415606131606362);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.366790615823915);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.350473827399395);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.220883466441019);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.136737383987297);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.185283200787522);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.154672366305157);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0558276893202577);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.179619522160829);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0788869523003634);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0133500996200616);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0134849491111732);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.00943946437782135);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0140243470756202);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0.00337123727779334);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.070389896781954 0
0.070389896781954 3
};
\addlegendentry{$\mu = 0.070$}
\addplot [semithick, black, dashed]
table {%
-0.131187651225067 1
0.236045424145365 1
};
\addlegendentry{$\sigma = 0.167$}
\end{axis}

\end{tikzpicture}