### 2021-07-23 latex_files/ch2/2_5_ROMC_algorithmic.tex

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI} that
follows. At a high level, ROMC can be split into the training and the
inference part. In general terms, the training part covers all the
appropriate steps for estimating the proposal regions and the
inference part for getting weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines formally ROMC;
Steps 2-10 (before the vertical line) form the training part, while
steps 12 and 13 the inference part.

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab)$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define deterministic distance function}
      \State Obtain $\thetab_i^*, d_i^*$ \Comment{Solve optimization problem}
    \EndFor
    \State $\text{L} \gets \{d_{i, i = 1, ..., n_1}: d_i^* < \epsilon \} $ \Comment{List with filtered solutions}
    \For{$i \gets 1 \textrm{ to } len(L)$}
      \State Construct $q_i$ \Comment{Compute proposal area}
      \State Compute $\Tilde{d}_i$ \Comment{Fit surrogate model in $q_i$}
      \EndFor
    \\\hrulefill
    \State Draw $\{w_{ij}, \thetab_{ij}\}$ \Comment{Draw $n_2$ samples per $q_i$}
    \State Draw $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}    
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are; (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ for obtaining the deterministic
functions $g_i(\thetab)$, (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$, (c) obtaining $\thetab_i^*$ and
$d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and (e) building
the bounding box for obtaining the proposal region $q_i$. If gradients
are available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
in closed-form, the procedure is upgraded in terms of both accuracy
and efficiency; If closed-form description is not available, gradients
approximation with finite-differences requires two evaluations of
$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
Bayesian Optimisation can be used as an alternative solution. In this
scenario, the training part becomes slower due to fitting of the
surrogate model and the blind optimisation
steps. Algorithm~\ref{alg:training_GB} defines formally the steps of a
gradient-based optimiser and algorithm~\ref{alg:training_GP} the steps
of the Bayesian Optimisation. After obtaning the optimal points
$\thetab^*$, the training part gets completed by estimating the
proposal regions. Algorithm~\ref{alg:region_construction} describes
the line search approach for finding the region's boundaries.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
expectation can be done easily after weighted samples are obtained
using the equation~\ref{eq:expectation}, so we do not discuss it
separately.

\noindent
Sampling is performed by getting $n_2$ samples from each proposal
distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
checking if it lies inside the acceptance region. If this is the case,
the corresponding weight is computed as in \eqref{eq:sampling}. As
before, if a surrogate model $\hat{d}$ is available, it can be
utilised for evaluating the indicator function.

\noindent
Evaluating the unnormalised posterior requires the deterministic
functions $g_i$ and the prior distribution $p(\thetab)$\footnote{There
  is no need for solving the optimisation problems and building the
  proposal regions.}. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data.

\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*$ using a Gradient Optimiser
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), g_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $H_i \approx \jac^T_i \jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}

\begin{algorithm}[!ht]
	\caption{Computation of the proposal distribution $q_i$; Needs, a model of distance $d$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\Repeat
        	\Repeat
                \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
    	\Until $k = K$
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \For {$i \gets 1 \textrm{ to } n_1$}
      \For {$j \gets 1 \textrm{ to } n_2$}
          \State $\thetab_{ij} \sim q_i$
          \If {$g_i(\thetab_{ij}) > \epsilon$}
            \State Reject $\theta_{ij}$
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
          \EndIf
      \EndFor
      \EndFor
    \end{algorithmic}
\end{algorithm}