### 2021-07-27 latex_files/ch2/2_1_LFI_intro.tex
minor improvements

An implicit or simulator-based\footnote{The terms can be used
  interchangeably} model is a parameterised stochastic data generating
mechanism \cite{Gutmann2016}. The key characteristic of these models
is that although it is feasible to sample (simulate) data points, it
is impossible to evaluate the likelihood. Formally, a simulator-based
model is a parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$ whose closed-form is either
unknown or intractable to evaluate. Practically, a simulator can be
understood as a black-box machine $M_r$\footnote{The subscript $r$ in
  $M_r$ indicates the \textit{random} simulator. In the next chapters
  we will introduce $M_d$ witch stands for the \textit{deterministic}
  simulator.} that given a set of parameters $\thetab$, produces
samples $\yb$ in a stochastic manner i.e.\
$M_r(\thetab) \rightarrow \yb$. Simulator-based models are
particularly captivating due to the modelling freedom they provide;
any physical process, that can be conceptualised as a computer program
of finite steps, can be represented as a simulator-based model. For
this reason, they are quite popular in many fiels such as biology,
neuroscience etc. This degree of freedom comes at a cost; performing
the inference is particularly demanding from both computational and
mathematical perspective.

\subsubsection*{\textit{Goal of implicit models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the parameters of
interest $\thetab$ as random variables and we try to \textit{infer} a
posterior distribution $p(\thetab|\data)$ on them. The class of
methods for approximate the posterior distribution in implicit models
is called likelihood-free inference methods. Unfortunately, the
algorithms deployed so far permit the inference only at
low-dimensional parametric spaces, i.e.\ $\thetab \in \mathbb{R}^D$
where $D$ is small\footnote{In broad terms, small refers to
  dimensionality $D$ less than a hundred.}.