### 2021-07-27 latex_files/ch2/2_1_implicit_models.tex
michael updates

%% Not used; Michael combined the text with the LFI intro

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic
of these models is that we can sample (simulate) data points,
but that we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models provide considerable modelling freedom; any physical process that can be
conceptualised as a computer program of finite steps can be modelled
as a simulator-based model without any compromise. The modelling
freedom allows for any amount of hidden (unobserved) internal variables
or rule-based decisions. Hence, implicit models are often used to
model physical phenomena in the natural sciences such as genetics, epidemiology, or neuroscience, to name a few examples.
The modelling freedom, however, makes parameter inference difficult.

% ;performing the inference
% is particularly demanding from both the computational and the
% mathematical perspective. The class of methods that try to approximate
% the posterior distribution in these cases are called likelihood-free
% inference methods. Unfortunately, the algorithms deployed so far
% permit the inference only at low-dimensional parametric spaces, i.e.\
% $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms,
%   small refers to dimensionality $D$ less than a hundred.}.

\subsubsection*{\textit{Goal of implicit models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the
parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.