### 2021-08-24 latex_files/images/chapter4/mae2_hist_t1_romc.tex

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - Histogram of the parameter \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.23969307221967);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.22260605418476);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.58099417689925);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.86697039022021);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.9327905830084);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.28560027755721);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.65528249842742);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.76542029517717);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.08006607906211);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.37496095644731);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.55021219168023);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.48979291667414);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.29497485978374);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.21369852560985);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.95263129250428);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.68186180204391);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.34177450679284);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.16665229531096);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.90550699050888);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.80681861350051);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.82068989944754);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.66308656575326);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.55350554347052);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.55706027108261);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.34606411506585);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.25058859381965);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.02541420681756);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.08883061316133);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.901026681844632);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.90959073604536);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.663388353950532);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.635304321110202);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.584894248497155);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.541909616027169);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.364723435432413);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.337243430980688);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.304888876342628);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.11082071864216);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.14353193766205);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.12562311229572);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.107714286929387);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0820282878821482);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0827149079075519);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0269285717323464);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0307869571533697);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.50403516578119 0
0.50403516578119 3.5
};
\addlegendentry{$\mu = 0.504$}
\addplot [semithick, black, dashed]
table {%
0.321847690178847 1
0.637029674539771 1
};
\addlegendentry{$\sigma = 0.143$}
\end{axis}

\end{tikzpicture}