### 2021-08-24 latex_files/images/chapter4/mae2_hist_t2_romc.tex

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - Histogram of the parameter \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00554428884192465);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0609051679887177);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.172932144280507);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.33874934889258);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.484235646321314);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.665208164482786);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.92463814505837);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.12139409849263);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.18629739177248);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.38911741112759);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.3586665597235);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.4168730451184);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.86440804043856);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.77193230571172);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.14800442101592);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.34373098482187);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24893542723874);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.36527262608105);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.32011000423491);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.35112752816106);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.09237672296862);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.72225524667604);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65885907591305);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.4089028988961);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18892677075576);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.11057927007035);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.808010795099994);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.746145310081448);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.68232985247479);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.586634734027065);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.400470451389202);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.259104715630191);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.230773630659991);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.211235556781049);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129869421833243);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0903839207491964);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0434117816322698);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0482353129247447);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0289411877548468);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0144705938774233);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.0325607299281219 0
0.0325607299281219 3
};
\addlegendentry{$\mu = 0.033$}
\addplot [semithick, black, dashed]
table {%
-0.17677281465975 1
0.198406420501618 1
};
\addlegendentry{$\sigma = 0.171$}
\end{axis}

\end{tikzpicture}