### 2021-08-27 latex_files/images/chapter3/example_post.tex
added Neural net

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.75,0.75,0}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Approximate Posterior},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta\)},
xmin=-3, xmax=3,
xtick style={color=black},
xtick={-3,-2,-1,0,1,2,3},
xticklabels={
  \(\displaystyle {\ensuremath{-}3}\),
  \(\displaystyle {\ensuremath{-}2}\),
  \(\displaystyle {\ensuremath{-}1}\),
  \(\displaystyle {0}\),
  \(\displaystyle {1}\),
  \(\displaystyle {2}\),
  \(\displaystyle {3}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=0.6,
ytick style={color=black},
ytick={0,0.1,0.2,0.3,0.4,0.5,0.6},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.1}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.3}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {0.6}\)
}
]
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-3,0) rectangle (axis cs:-2.88,0);
\addlegendimage{ybar,ybar legend,draw=none,fill=color0,fill opacity=0.5}
\addlegendentry{samples histogram}

\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.88,0) rectangle (axis cs:-2.76,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.76,0) rectangle (axis cs:-2.64,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.64,0) rectangle (axis cs:-2.52,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.52,0) rectangle (axis cs:-2.4,0.0232073945389685);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.4,0) rectangle (axis cs:-2.28,0.0370673662775194);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.28,0) rectangle (axis cs:-2.16,0.0432283769706026);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.16,0) rectangle (axis cs:-2.04,0.0612233170218504);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.04,0) rectangle (axis cs:-1.92,0.0863738703892799);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.92,0) rectangle (axis cs:-1.8,0.109332614272474);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.8,0) rectangle (axis cs:-1.68,0.1270696943844);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.68,0) rectangle (axis cs:-1.56,0.162608319593082);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.56,0) rectangle (axis cs:-1.44,0.165932870953625);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.44,0) rectangle (axis cs:-1.32,0.186184085473939);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.32,0) rectangle (axis cs:-1.2,0.21004533914476);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.2,0) rectangle (axis cs:-1.08,0.260281980894789);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.08,0) rectangle (axis cs:-0.96,0.258872960512065);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.96,0) rectangle (axis cs:-0.84,0.267861221254142);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.84,0) rectangle (axis cs:-0.72,0.301309339097609);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.72,0) rectangle (axis cs:-0.6,0.296677069473362);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.6,0) rectangle (axis cs:-0.48,0.318079444437077);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.48,0) rectangle (axis cs:-0.36,0.310739645449952);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.36,0) rectangle (axis cs:-0.24,0.317600561692622);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.24,0) rectangle (axis cs:-0.12,0.313833964721813);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.12,0) rectangle (axis cs:0,0.340218562084569);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0,0) rectangle (axis cs:0.12,0.343469439176735);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.12,0) rectangle (axis cs:0.24,0.315528472894501);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.24,0) rectangle (axis cs:0.36,0.327491332222326);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.36,0) rectangle (axis cs:0.48,0.320667253113843);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.48,0) rectangle (axis cs:0.6,0.309533229305267);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.6,0) rectangle (axis cs:0.72,0.307884767550316);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.72,0) rectangle (axis cs:0.84,0.289208340516575);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.84,0) rectangle (axis cs:0.96,0.272217212371973);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.96,0) rectangle (axis cs:1.08,0.235895798061777);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.08,0) rectangle (axis cs:1.2,0.221437222892658);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.2,0) rectangle (axis cs:1.32,0.207761436825051);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.32,0) rectangle (axis cs:1.44,0.206149812204291);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.44,0) rectangle (axis cs:1.56,0.191175517156526);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.56,0) rectangle (axis cs:1.68,0.130919174907133);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.68,0) rectangle (axis cs:1.8,0.118127480060059);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.8,0) rectangle (axis cs:1.92,0.0924796253810799);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.92,0) rectangle (axis cs:2.04,0.0812258808863897);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.04,0) rectangle (axis cs:2.16,0.0618587575866078);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.16,0) rectangle (axis cs:2.28,0.0384119216754122);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.28,0) rectangle (axis cs:2.4,0.0377120161258239);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.4,0) rectangle (axis cs:2.52,0.026430643780492);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.52,0) rectangle (axis cs:2.64,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.64,0) rectangle (axis cs:2.76,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.76,0) rectangle (axis cs:2.88,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.88,0) rectangle (axis cs:3,0);
\addplot [semithick, blue, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.51677846908569 0
-2.4765100479126 0.200000047683716
2.4765100479126 0.200000047683716
2.51677846908569 0
3 0
};
\addlegendentry{Prior}
\addplot [semithick, green!50!black, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0.0149635076522827
-2.91946315765381 0.0183337926864624
-2.83892607688904 0.0223180055618286
-2.75838923454285 0.0269923210144043
-2.67785239219666 0.0324347019195557
-2.63758397102356 0.0354681015014648
-2.59731554985046 0.0387222766876221
-2.55704689025879 0.0422066450119019
-2.51677846908569 0.0459299087524414
-2.4765100479126 0.0499007701873779
-2.4362416267395 0.0541269779205322
-2.39597320556641 0.058616042137146
-2.35570478439331 0.0633745193481445
-2.31543612480164 0.0684083700180054
-2.27516770362854 0.073722243309021
-2.23489928245544 0.0793203115463257
-2.19463086128235 0.0852051973342896
-2.15436244010925 0.0913783311843872
-2.11409401893616 0.0978399515151978
-2.07382559776306 0.104588747024536
-2.03355693817139 0.111621975898743
-1.99328863620758 0.118935108184814
-1.9530200958252 0.126521944999695
-1.9127516746521 0.134374856948853
-1.872483253479 0.142483830451965
-1.83221471309662 0.150837302207947
-1.79194629192352 0.159421920776367
-1.71140944957733 0.177220344543457
-1.63087248802185 0.195732235908508
-1.55033552646637 0.214780211448669
-1.3892617225647 0.253633737564087
-1.30872488021851 0.272953271865845
-1.22818791866302 0.291845083236694
-1.14765095710754 0.310027003288269
-1.10738253593445 0.318762183189392
-1.06711411476135 0.327212452888489
-1.02684569358826 0.335342526435852
-0.986577153205872 0.343117833137512
-0.946308732032776 0.350504517555237
-0.906040191650391 0.357470154762268
-0.865771770477295 0.363983392715454
-0.825503349304199 0.370014905929565
-0.785234928131104 0.375536918640137
-0.744966506958008 0.380523920059204
-0.704697966575623 0.384952306747437
-0.664429545402527 0.388801217079163
-0.624161005020142 0.392052412033081
-0.583892583847046 0.394690275192261
-0.54362416267395 0.39670205116272
-0.503355741500854 0.398078083992004
-0.463087320327759 0.398520588874817
-0.382550358772278 0.398850798606873
-0.18120801448822 0.398941993713379
0.422818779945374 0.398738622665405
0.463087320327759 0.398520588874817
0.503355741500854 0.398078083992004
0.54362416267395 0.39670205116272
0.583892583847046 0.394690275192261
0.624161005020142 0.392052412033081
0.664429545402527 0.388801217079163
0.704697966575623 0.384952306747437
0.744966506958008 0.380523920059204
0.785234928131104 0.375536918640137
0.825503349304199 0.370014905929565
0.865771770477295 0.363983392715454
0.906040191650391 0.357470154762268
0.946308732032776 0.350504517555237
0.986577153205872 0.343117833137512
1.02684569358826 0.335342526435852
1.06711411476135 0.327212452888489
1.10738253593445 0.318762183189392
1.14765095710754 0.310027003288269
1.22818791866302 0.291845083236694
1.30872488021851 0.272953271865845
1.42953014373779 0.24389910697937
1.55033552646637 0.214780211448669
1.63087248802185 0.195732235908508
1.71140944957733 0.177220344543457
1.79194629192352 0.159421920776367
1.83221471309662 0.150837302207947
1.872483253479 0.142483830451965
1.9127516746521 0.134374856948853
1.9530200958252 0.126521944999695
1.99328863620758 0.118935108184814
2.03355693817139 0.111621975898743
2.07382559776306 0.104588747024536
2.11409401893616 0.0978399515151978
2.15436244010925 0.0913783311843872
2.19463086128235 0.0852051973342896
2.23489928245544 0.0793203115463257
2.27516770362854 0.073722243309021
2.31543612480164 0.0684083700180054
2.35570478439331 0.0633745193481445
2.39597320556641 0.058616042137146
2.4362416267395 0.0541269779205322
2.4765100479126 0.0499007701873779
2.51677846908569 0.0459299087524414
2.55704689025879 0.0422066450119019
2.59731554985046 0.0387222766876221
2.63758397102356 0.0354681015014648
2.67785239219666 0.0324347019195557
2.75838923454285 0.0269923210144043
2.83892607688904 0.0223180055618286
2.91946315765381 0.0183337926864624
3 0.0149635076522827
};
\addlegendentry{Likelihood}
\addplot [semithick, blue!50.1960784313725!black, dash pattern=on 1pt off 3pt on 3pt off 3pt, mark=*, mark size=3, mark options={solid}]
table {%
-3 0
-2.9597315788269 0
-2.91946315765381 0
-2.87919473648071 0
-2.83892607688904 0
-2.79865765571594 0
-2.75838923454285 0
-2.71812081336975 0
-2.67785239219666 0
-2.63758397102356 0
-2.59731554985046 0
-2.55704689025879 0
-2.51677846908569 0
-2.4765100479126 0.0550098419189453
-2.4362416267395 0.0559921264648438
-2.39597320556641 0.060903787612915
-2.35570478439331 0.0648330450057983
-2.31543612480164 0.0658153295516968
-2.27516770362854 0.0697445869445801
-2.23489928245544 0.0746562480926514
-2.19463086128235 0.0834970474243164
-2.15436244010925 0.0903732776641846
-2.11409401893616 0.0982317924499512
-2.07382559776306 0.105108022689819
-2.03355693817139 0.111984252929688
-1.99328863620758 0.116895914077759
-1.9530200958252 0.123772144317627
-1.9127516746521 0.126718997955322
-1.872483253479 0.131630659103394
-1.83221471309662 0.13948917388916
-1.79194629192352 0.14538311958313
-1.75167787075043 0.149312376976013
-1.71140944957733 0.154223918914795
-1.67114090919495 0.160117864608765
-1.63087248802185 0.168958783149719
-1.59060406684875 0.176817297935486
-1.55033552646637 0.181728839874268
-1.51006710529327 0.190569758415222
-1.46979868412018 0.196463704109192
-1.42953014373779 0.199410557746887
-1.3892617225647 0.20923376083374
-1.3489933013916 0.214145421981812
-1.30872488021851 0.217092275619507
-1.26845633983612 0.224950909614563
-1.22818791866302 0.230844736099243
-1.18791949748993 0.235756397247314
-1.14765095710754 0.236738681793213
-1.10738253593445 0.245579600334167
-1.06711411476135 0.246561884880066
-1.02684569358826 0.250491142272949
-0.986577153205872 0.255402803421021
-0.946308732032776 0.259332060813904
-0.906040191650391 0.26719057559967
-0.865771770477295 0.27308452129364
-0.825503349304199 0.275049090385437
-0.785234928131104 0.281925320625305
-0.744966506958008 0.28487229347229
-0.704697966575623 0.288801550865173
-0.664429545402527 0.288801550865173
-0.624161005020142 0.287819266319275
-0.583892583847046 0.29076623916626
-0.54362416267395 0.292730808258057
-0.503355741500854 0.29076623916626
-0.463087320327759 0.29076623916626
-0.422818779945374 0.289783954620361
-0.382550358772278 0.289783954620361
-0.342281818389893 0.288801550865173
-0.302013397216797 0.288801550865173
-0.261744976043701 0.289783954620361
-0.221476554870605 0.287819266319275
-0.18120801448822 0.288801550865173
-0.140939593315125 0.288801550865173
-0.100671172142029 0.288801550865173
-0.0604026317596436 0.288801550865173
-0.0201342105865479 0.288801550865173
0.0201342105865479 0.288801550865173
0.0604026317596436 0.288801550865173
0.100671172142029 0.288801550865173
0.140939593315125 0.288801550865173
0.18120801448822 0.288801550865173
0.221476554870605 0.287819266319275
0.261744976043701 0.289783954620361
0.302013397216797 0.288801550865173
0.342281818389893 0.288801550865173
0.382550358772278 0.289783954620361
0.422818779945374 0.289783954620361
0.463087320327759 0.29076623916626
0.503355741500854 0.29076623916626
0.54362416267395 0.292730808258057
0.583892583847046 0.29076623916626
0.624161005020142 0.287819266319275
0.664429545402527 0.288801550865173
0.704697966575623 0.288801550865173
0.744966506958008 0.28487229347229
0.785234928131104 0.281925320625305
0.825503349304199 0.275049090385437
0.865771770477295 0.27308452129364
0.906040191650391 0.26719057559967
0.946308732032776 0.259332060813904
0.986577153205872 0.255402803421021
1.02684569358826 0.250491142272949
1.06711411476135 0.246561884880066
1.10738253593445 0.245579600334167
1.14765095710754 0.236738681793213
1.18791949748993 0.235756397247314
1.22818791866302 0.230844736099243
1.26845633983612 0.224950909614563
1.30872488021851 0.217092275619507
1.3489933013916 0.214145421981812
1.3892617225647 0.20923376083374
1.42953014373779 0.199410557746887
1.46979868412018 0.196463704109192
1.51006710529327 0.190569758415222
1.55033552646637 0.181728839874268
1.59060406684875 0.176817297935486
1.63087248802185 0.168958783149719
1.67114090919495 0.160117864608765
1.71140944957733 0.154223918914795
1.75167787075043 0.149312376976013
1.79194629192352 0.14538311958313
1.83221471309662 0.13948917388916
1.872483253479 0.131630659103394
1.9127516746521 0.126718997955322
1.9530200958252 0.123772144317627
1.99328863620758 0.116895914077759
2.03355693817139 0.111984252929688
2.07382559776306 0.105108022689819
2.11409401893616 0.0982317924499512
2.15436244010925 0.0903732776641846
2.19463086128235 0.0834970474243164
2.23489928245544 0.0746562480926514
2.27516770362854 0.0697445869445801
2.31543612480164 0.0658153295516968
2.35570478439331 0.0648330450057983
2.39597320556641 0.060903787612915
2.4362416267395 0.0559921264648438
2.4765100479126 0.0550098419189453
2.51677846908569 0
2.55704689025879 0
2.59731554985046 0
2.63758397102356 0
2.67785239219666 0
2.71812081336975 0
2.75838923454285 0
2.79865765571594 0
2.83892607688904 0
2.87919473648071 0
2.91946315765381 0
2.9597315788269 0
3 0
};
\addlegendentry{ROMC Posterior}
\addplot [semithick, red, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.51677846908569 0
-2.4765100479126 0.0380961894989014
-2.39597320556641 0.0447498559951782
-2.35570478439331 0.0483826398849487
-2.31543612480164 0.0522257089614868
-2.27516770362854 0.0562825202941895
-2.23489928245544 0.0605562925338745
-2.19463086128235 0.0650490522384644
-2.15436244010925 0.0697618722915649
-2.11409401893616 0.0746949911117554
-2.07382559776306 0.0798472166061401
-2.03355693817139 0.0852166414260864
-1.9530200958252 0.0965919494628906
-1.872483253479 0.10877788066864
-1.79194629192352 0.121709108352661
-1.71140944957733 0.135297179222107
-1.63087248802185 0.14942991733551
-1.55033552646637 0.163971900939941
-1.22818791866302 0.222806215286255
-1.14765095710754 0.236687064170837
-1.10738253593445 0.243355870246887
-1.06711411476135 0.249807119369507
-1.02684569358826 0.256013989448547
-0.986577153205872 0.261949896812439
-0.946308732032776 0.267589211463928
-0.906040191650391 0.272907018661499
-0.865771770477295 0.277879595756531
-0.825503349304199 0.282484292984009
-0.785234928131104 0.286700010299683
-0.744966506958008 0.290507197380066
-0.704697966575623 0.293888092041016
-0.664429545402527 0.296826481819153
-0.624161005020142 0.29930853843689
-0.583892583847046 0.301322460174561
-0.54362416267395 0.302858352661133
-0.503355741500854 0.303908824920654
-0.422818779945374 0.304413080215454
-0.261744976043701 0.304565191268921
0.422818779945374 0.304413080215454
0.503355741500854 0.303908824920654
0.54362416267395 0.302858352661133
0.583892583847046 0.301322460174561
0.624161005020142 0.29930853843689
0.664429545402527 0.296826481819153
0.704697966575623 0.293888092041016
0.744966506958008 0.290507197380066
0.785234928131104 0.286700010299683
0.825503349304199 0.282484292984009
0.865771770477295 0.277879595756531
0.906040191650391 0.272907018661499
0.946308732032776 0.267589211463928
0.986577153205872 0.261949896812439
1.02684569358826 0.256013989448547
1.06711411476135 0.249807119369507
1.10738253593445 0.243355870246887
1.18791949748993 0.229828119277954
1.26845633983612 0.215649008750916
1.3489933013916 0.20103645324707
1.63087248802185 0.14942991733551
1.71140944957733 0.135297179222107
1.79194629192352 0.121709108352661
1.872483253479 0.10877788066864
1.9530200958252 0.0965919494628906
2.03355693817139 0.0852166414260864
2.07382559776306 0.0798472166061401
2.11409401893616 0.0746949911117554
2.15436244010925 0.0697618722915649
2.19463086128235 0.0650490522384644
2.23489928245544 0.0605562925338745
2.27516770362854 0.0562825202941895
2.31543612480164 0.0522257089614868
2.35570478439331 0.0483826398849487
2.4362416267395 0.0413227081298828
2.4765100479126 0.0380961894989014
2.51677846908569 0
3 0
};
\addlegendentry{True Posterior}
\end{axis}

\end{tikzpicture}