### 2021-10-05 latex_files/ch4/4-4-extensibility.tex

ROMC can be conceived as a meta-algorithm; it describes a sequence of
steps for approximating the posterior distribution, without
\emph{explicitly} enforcing the methods that solve these steps. Even
though for each step a specific algotihm is proposed
by~\cite{Ikonomov2019}, in general ROMC works well if a practitioner
thinks of an alternative way of approaching a specific
step. Considering this particularity, we designed the implementation
to support such \emph{extensibility} capability.

We have specified four specific points where a user may intervene with
their custom modules; (a) the gradient-based optimization, (b) the
Bayesian Optimization, (c) the proposal region construction and (d)
the surrogate model fitting. These are the four critical parts of the
ROMC procedure, whereas the rest of the code is the backbone of the
algorithm; each of them may be approached with a completely different
algorithm than the one we propose, without the rest of the method to
collapse.

The four replaceable parts described above, are solved using the four
methods of the \linebreak \code{OptimisatioProblem} class; (a)
\code{solve_gradients(**kwargs)}, (b) \code{solve_bo(**kwargs)},
\linebreak (c) \code{build_region(**kwargs)}, (d)
\code{fit_local_surrogate(**kwargs)}. Therefore, the user should
create a custom class that inherits the basic
\code{OptimisatioProblem} class and overwrites one or more of the
above four functions with custom ones.

We illustrate that in the
following example; suppose we want to blindly set the proposal region
as a predefined n-dimensional bounding box around the optimal
point. The following snippet illustrates how:

\begin{Code}
---------------------------------- python ----------------------------------
class customOptim(OptimisationProblem):
    def __init__(self, **kwargs):
        super(customOptim, self).__init__(**kwargs)

    def fit_local_surrogate(self, **kwargs):
        nof_samples = 500
        objective = self.objective

        local_surrogates = []
        for i in range(len(self.regions)):
            # prepare dataset
            x = self.regions[i].sample(nof_samples)
            y = np.array([objective(ii) for ii in x])

            # train Neural Network
            mlp = MLPRegressor(hidden_layer_sizes=(10,10), solver='adam')
            model = Pipeline([('linear', mlp)])
            model = model.fit(x, y)
            local_surrogates.append(self.create_local_surrogate(model))

        self.local_surrogates = local_surrogates
        self.state["local_surrogates"] = True

    def create_local_surrogate(self, model):
        def _local_surrogate(th):
            th = np.expand_dims(th, 0)
            return float(model.predict(th))
        return _local_surrogate
----------------------------------------------------------------------------    
\end{Code}

In the same way, the user may replace each of the other three
functionalities. The only restriction that must be respected, concerns
the side-effects each method has at the \code{OptimisatioProblem}
class level. In the following four snippets we present the signature
of each method, pointing out with a left arrow the side-effects that
must be included in the custom routine that will replace them.

\begin{Code}
---------------------------------- python ----------------------------------
def solve_gradients(self, **kwargs):
  self.state["attempted"] = True

  # custom solution procedure

  if success:
    self.result = RomcOpimisationResult(res.x, res.y, jac, hess_inv)
    self.state["solved"] = True
    return True 
  else:
    return False 
----------------------------------------------------------------------------    
def solve_bo(self, **kwargs):
  self.state["attempted"] = True

  # custom procedure

  if success:
    self.result = RomcOpimisationResult(x, y)
    self.surrogate <- Callable
    self.state["solved"] = True
    self.state["has_fit_surrogate"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------    
def build_region(self, **kwargs):
    self.eps_region = eps_region

    if success:
        # construct region
        self.regions <- List[NDimBoudningBox]
        self.state["region"] = True
        return True
    else:
        return False
----------------------------------------------------------------------------    
def fit_local_surrogate(self, **kwargs):
  if success:
    self.local_surrogate = local_surrogates
    self.state["local_surrogates"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------    
\end{Code}

The two classes that may be needed for creating the custom routines
are \\ (a) \code{RomcOpimisationResult} and (b)
\code{NDimBoundingBox}. We present their signatures below.

\begin{Code}
---------------------------------- python ----------------------------------
class RomcOpimisationResult:
    def __init__(self, x_min, f_min, hess_appr):
        Parameters
        ----------
        x_min: np.ndarray (D,) or float, the minimum point
        f_min: float, distance at the minimum point
        hess_appr: np.ndarray (DxD), Hessian approximation at x_min
        """
----------------------------------------------------------------------------    
class NDimBoundingBox:
    def __init__(self, rotation, center, limits, eps_region):
        Parameters
        ----------
        rotation: np.array (D,D),  rotation matrix for the bounding box
        center: np.array (D,) center of the bounding box
        limits: np.ndarray, shape: (D,2), the limits of the bounding box
        eps_region: float, distance threshold 
----------------------------------------------------------------------------    
\end{Code}

% Let's say we have observed that the local area around $\theta_i^*$ is
% too complex to be represented by a simple quadratic model\footnote{as
%   in the current implementation}. Hence, the user selects a neural
% network as a good alternative. In the following snippet, we
% demonstrate how they could implement this enhancement without much
% effort; (a) they have to develop the neural network using the package
% of their choice (b) they must create a custom optimisation class which
% inherits the basic \code{OptimisationClass} and (c) they have to
% overwrite the \code{fit_local_surrogate} routine, with one that
% sets the neural network's prediction function as the
% \code{local_surrogate} attribute. The argument \code{**kwargs}
% may be used for passing all the important arguments, e.g.\ training
% epochs, gradient step etc. If, for example, they would like to set the
% size of the training set dynamically, we may replace \code{x =
%   self.regions[0].sample(30)} with \code{x =
%   self.regions[0].sample(kwargs["nof_examples"])}. Finally, they must
% pass the custom optimisation class, when calling the \code{ROMC}
% method.

% \begin{Code}
% ------------------------------ python snippet ------------------------------  
%   class NeuralNetwork:
%       def __init__(self, **kwargs):
%           # set the input arguments

%       def train(x, y):
%           # training code

%       def predict(x):
%           # prediction code

%   # Inherit the base optimisation class
%   class customOptim(elfi.methods.parameter_inference.OptimisationProblem):
%       def __init__(self, **kwargs):
%           super(customOptim, self).__init__(**kwargs)

%       # overwrite the function you want to replace
%       def fit_local_surrogate(**kwargs):
%           # init and train the NN
%           x = self.regions[0].sample(30) # 30 training points
%           y = [np.array([self.objective(ii) for ii in x])]
%           nn = NeuralNet()
%           nn.train(x,y)

%           # set the appropriate attribute
%           self.local_surrogate = nn.predict

%           # update the state
%           self.state["local_surrogate"] = True

%   # pass the custom inference method as argument
%   romc = elfi.ROMC(dist, bounds, custom_optim_class=customOptim)
% ----------------------------------------------------------------------------
% \end{Code}