### 2022-01-27 paper.tex

\documentclass[article]{jss}
%% -- LaTeX packages and custom commands ---------------------------------------

\usepackage[usenames,dvipsnames]{xcolor}

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}

\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption} % for subfigure
\usepackage{float}

%% TikZ packages
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots}
\pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}

%% for algorithms
\usepackage{algorithm}
\usepackage{algpseudocode} % algorithmicx package

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{marginnote}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% Vasilis' custom commands
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\X}{\mathbf{x}}
\newcommand{\Z}{\mathbf{z}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\hess}{\mathbf{H}}
\newcommand{\Thetab}{\mathbf{\Theta}}

\newcommand{\vb}{\mathbf{v}}
\newcommand{\yb}{\mathbf{y}}
\newcommand{\xb}{\mathbf{x}}

\newcommand{\jac}{\mathbf{J}}
\newcommand{\hessian}{\mathbf{H}}
\newcommand{\thetab}{\boldsymbol{\theta}}
\newcommand{\thetabij}{\thetab_{ij}}
\newcommand{\thetabi}{\thetab_i}

\newcommand{\simulator}{g}
\newcommand{\region}{B_{d,\epsilon}}
\newcommand{\indicator}[1]{\mathbb{1}_{#1}}
\newcommand{\regioni}{B_{d,\epsilon}^i}
\newcommand{\data}{\mathbf{y_0}}
\newcommand{\accregioni}{C^i_{\epsilon}}
\newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}

\newcommand{\danger}{\fontencoding{U}\fontfamily{futs}\selectfont\char 66\relax}
\newcommand{\TODO}[1]{\textcolor{red}{TODO #1}\marginpar{\textcolor{red}{\Large \danger}}}

\newcommand{\R}{\mathbb{R}}

\newcommand{\Ex}{\mathbb{E}}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Vasilis Gkolemis\\ATHENA RC
  \And Michael Gutmann\\University of Edinburgh
   \And Henri Pesonen \\University of Oslo
}
\Plainauthor{Vasilis Gkolemis, Michael Gutmann}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}

%% - \Abstract{} almost as usual
\Abstract{
  Likelihood-free inference (LFI) or Approximate Bayesian Computation (ABC) is the general framework for performing the inference in statistical models with intractable likelihood. In general, the intractability of the likelihood makes the inference problem quite challenging, therefore, many LFI methods either lack accuracy or are pretty inefficient. Robust Optimisation Monte Carlo (ROMC) is a highly-parallelizable approximate Bayesian computation (ABC) technique. In this paper, we present the implementation of ROMC in the \proglang{Python} package ELFI. Our implementation can be exploited in two directions. A scientist may use it as an out-of-the-box ABC algorithm for performing the inference in an LFI problem; we provide an easy-to-use API harmonized with the rest of the ELFI framework. Therefore, it is directly comparable with the rest of the methods included in ELFI. On the other direction, we have carefully designed ROMC in isolated algorithmic steps. Therefore, we allow the researcher to experiment with ROMC, altering or replacing parts of the algorithm with custom-created functionalities. In both scenarios, the algorithmic steps of ROMC can run in a fully-parallelizable manner, exploiting all CPU cores. We also provide some functionalities for visualization and evaluation needs. Finally, we test the robustness of our implementation on some typical LFI models that ship.

  % This short article illustrates how to write a manuscript for the
  % \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  % Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  % it is recommended to keep the {\LaTeX} code as simple as possible,
  % i.e., avoid inclusion of packages/commands that are not necessary.
  % For outlining the typical structure of a JSS article some brief text snippets
  % are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  % discussing count data regression in \proglang{R}. Editorial comments and
  % instructions are marked by vertical bars.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Bayesian Inference, Simulator-based models, Likelihood-free, \proglang{Python}, \proglang{ELFI}}
\Plainkeywords{Bayesian Inference, Simulator-based models, Likelihood-free, Python, ELFI}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Vasilis Gkolemis\\
  Information Management Systems Institute (IMSI)\\
  ATHENA Research and Innovation Center\\
  Athens, Greece\\
  E-mail: \email{vgkolemis@athenarc.gr}\\
  URL: \url{https://givasile.github.io}
}

\begin{document}

%% See jss_template_article.tex for formatting instructions and macros to e.g. typset code
%% I only kept the informative sidebars as comments

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
\section{Introduction} \label{sec:intro}

Simulator-based models are particularly captivating due to the modeling freedom they provide. In essence, a simulator-based model can describe any data generating mechanism that can be written as a finite set of algorithmic steps. This modeling freedom comes at a cost; performing the inference, i.e., sample or evaluate the posterior distribution, is challenging.

Optimization Monte Carlo (OMC) proposed by [..] was a novel ABC approach approximating the posterior distribution. Afterward, Forneron and Ng provided a similar method under "reverse sampler ". The key idea is turning the stochastic data-generating mechanism into a set of fully-parallelizable deterministic optimization processes. Robust OMC (ROCM) provided an upgrade in OMC by locating and resolving some fundamental failure modes of the latter.

In this paper, we present an extendable and parallelizable implementation of ROMC at the Python package 'Engine for Likelihood-Free inference (ELFI) .' We have followed specific design principles for making the implementation extendible. Each step of the method is designed as a fully-isolated and parallelizable step; The user may easily replace parts of the technique with custom components without the backbone method collapse. Therefore, our implementation is a friendly tool for the research community to experiment with alterations of ROMC. Apart from the main functionalities, we also provide essential visualization and evaluation tools for gaining insights while fitting data. Finally, we verify the robustness of the implementation on the LFI examples supplied by the ELFI package.

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)

% For writing about software JSS requires authors to use the markup
% \verb|\proglang{}| (programming languages and large programmable systems),
% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
% arguments, etc.). If there is such markup in (sub)section titles (as above), a
% plain text version has to be provided in the {\LaTeX} command as well. Below we
% also illustrate how abbrevations should be introduced and citation commands can
% be employed. See the {\LaTeX} code for more details.
% \end{leftbar}

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

% \begin{leftbar}
% As the synopsis above is a code listing that is not meant to be executed,
% one can use either the dedicated \verb|{Code}| environment or a simple
% \verb|{verbatim}| environment for this. Again, spaces before and after should be
% avoided.

% Finally, there might be a reference to a \verb|{table}| such as
% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
% column headers and captions in sentence style, and if possible avoiding vertical
% lines.
% \end{leftbar}

%% -- Illustrations ------------------------------------------------------------

% - Virtually all JSS manuscripts list source code along with the generated
%   output. The style files provide dedicated environments for this.
% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%   or knitr using the render_sweave() hook - are used (without the need to
%   load Sweave.sty).
% - Equivalently, {CodeInput} and {CodeOutput} can be used.
% - The code input should use "the usual" command prompt in the respective
%   software system.
% - For R code, the prompt "R> " should be used with "+  " as the
%   continuation prompt.
% - Comments within the code chunks should be avoided - these should be made
%   within the regular LaTeX text.

% \begin{leftbar}
% For code input and output, the style files provide dedicated environments.
% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
% hook. Please make sure that all code is properly spaced, e.g., using
% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
% use ``the usual'' command prompt in the respective software system. For
% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
% the continuation prompt. Generally, comments within the code chunks should be
% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
% before and after code input/output should be avoided (see above).
% \end{leftbar}

% \clearpage
\section{Background}
We first give a short introduction to implicit parametric models and
likelihood-free inference. We then focus on the methods relevant for
this paper: first the earlier Optmisation Monte Carlo (OMC) and then
its recent robust generalisation, robust OMC, for which we provide the
software package as part of the Engine for likelihood-free inference
(ELFI), a larger inference platform that we briefly introduce at the
end of this background section.

%\input{latex_files/ch2/2_1_implicit_models}

\subsection{Implicit models and likelihood-free inference}
\input{latex_files/ch2/2-1-implicit-models.tex}

\subsection{Optimisation Monte Carlo (OMC)}
\input{latex_files/ch2/2-2-OMC.tex}

\subsection{Robust optimisation Monte Carlo (ROMC)}
\input{latex_files/ch2/2-3-ROMC.tex}

\subsection{Engine for likelihood-free inference (ELFI)}
\label{subsec:ELFI}
\input{latex_files/ch2/2-4-ELFI.tex}

%% -- Design Principles -----------------------------------------------------
\section{Design principles}  % better title needed

\input{latex_files/ch3/3-0-design-principles.tex}

\subsection{General implementation principles}
\label{subsec:general_design}
\input{latex_files/ch3/3-1-implementation-principles.tex}

%% -- ROMC implementation ----------------------------------------------------
\section{Implemented functionalities}

In this section, we present the implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{Engine
  for likelihood-free inference
  (ELFI)}. At~\ref{subsec:general_design} we explain the general
design that we follow. Then, we divide the implementation into three
parts; at~\ref{subsec:training} we present the training (fitting)
part, at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

\subsection{Training part} 
\label{subsec:training}
\input{latex_files/ch4/4-1-training.tex}

\subsection{Inference part} 
\label{subsec:inference}
\input{latex_files/ch4/4-2-inference.tex}

\subsection{Evaluation part} 
\label{subsec:evaluation}
\input{latex_files/ch4/4-3-evaluation.tex}

\subsection{Extensibility} % better title
\label{subsec:extensibility}
\input{latex_files/ch4/4-4-extensibility.tex}

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}

In this section, we illustrate the implementation performing two
experiments; In chapter~\ref{subsec:ma2} we perform the inference at
the second-order moving average (MA2) model, which is a standard model
of \pkg{ELFI}, and in chapter~\ref{subsec:exec_time} we measure the
execution speed-up achieved by the parallelization of ROMC.

\subsection{Moving average example}
\label{subsec:ma2}
\input{latex_files/ch5/5-1-ma2.tex}

\subsection{Execution time}
\label{subsec:exec_time}
\input{latex_files/ch5/5-2-execution-time.tex}

%% -- Summary/conclusions/discussion -------------------------------------------
\section{Summary and discussion} \label{sec:summary}

%% -- Optional special unnumbered sections -------------------------------------

% \section*{Computational details}

% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

% The results in this paper were obtained using
% \proglang{R}~3.4.1 with the
% \pkg{MASS}~7.3.47 package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

% \section*{Acknowledgments}

% \begin{leftbar}
% All acknowledgments (note the AE spelling) should be collected in this
% unnumbered section before the references. It may contain the usual information
% about funding and feedback from colleagues/reviewers/etc. Furthermore,
% information such as relative contributions of the authors may be added here
% (if any).
% \end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

% \newpage

% \begin{appendix}

% \section{More technical details} \label{app:technical}

% \begin{leftbar}
% Appendices can be included after the bibliography (with a page break). Each
% section within the appendix should have a proper section title (rather than
% just \emph{Appendix}).

% For more technical style details, please check out JSS's style FAQ at
% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
% which includes the following topics:
% \begin{itemize}
%   \item Title vs.\ sentence case.
%   \item Graphics formatting.
%   \item Naming conventions.
%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
%   \item Trouble shooting.
%   \item Many other potentially helpful details\dots
% \end{itemize}
% \end{leftbar}

% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

% \begin{leftbar}
% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
% formats of author-year citations and allow to include additional details (e.g.,
% pages, chapters, \dots) in brackets. In case you are not familiar with these
% commands see the JSS style FAQ for details.

% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
% when acquiring the entries automatically from mixed online sources. However,
% it is important that informations are complete and presented in a consistent
% style to avoid confusions. JSS requires the following format.
% \begin{itemize}
%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
%     be used in the references.
%   \item Titles should be in title case.
%   \item Journal titles should not be abbreviated and in title case.
%   \item DOIs should be included where available.
%   \item Software should be properly cited as well. For \proglang{R} packages
%     \code{citation("pkgname")} typically provides a good starting point.
% \end{itemize}
% \end{leftbar}

% \end{appendix}

%% -----------------------------------------------------------------------------

\end{document}