### 2022-02-02 latex_files/ch2/2-1-implicit-models.tex
Add comments and some changes. Ask HP.

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample, or simulate, data points, but that we cannot
evaluate the likelihood. Formally, a simulator-based model is a
parameterised family of probability density functions $\{
p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either unknown or
computationally intractable. Whereas evaluating $p(\yb|\thetab)$ is
intractable, sampling is feasible. Practically, a simulator can be
understood as a random black-box machine (computer code) that
generates samples $\yb$ in a stochastic manner for any given a set of
parameters $\thetab$. We can isolate the randomness by introducing the
stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator
becomes a deterministic mapping $\simulator$ that maps
$(\thetab, \vb)$ to the data $\yb$, i.e.\
$\yb=\simulator(\thetab,\vb)$. The distribution $p(\vb)$ is determined
by the random draws performed when running the computer
code. \textbf{HP comment: I don't think that the
distribution $p(\vb)$ is determined by the random draws, but $p(\vb)$ 
should define the way we choose the `random' draws. How about: The distribution 
$p(\vb)$ is defined as a random number generating process within the computer 
code.} Importantly, we can fix the value of $\vb$ while changing
$\thetab$ by fixing the random seed used.

Simulator-based models provide considerable modelling freedom; any
physical process that can be conceptualised as a computer program of
finite steps can be modelled as a simulator-based model without any
compromise. The modelling freedom allows for any amount of hidden
(unobserved) internal variables or rule-based decisions. Hence,
implicit models are often used to model physical phenomena in the
natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
Further background on simulator-based models and
example applications can be found in the articles
by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. Denoting observed
data by $\data$, the main difficulty is that the likelihood function
$L(\thetab) = p(\data|\thetab)$ is generally intractable. To better see the
sources of the intractability, and to address them, we go back to the
basic characterisation of the likelihood as the (rescaled) probability
of a parameter of the model to generate data $\yb$ that is similar to
the observed data $\data$. More formally, the likelihood $L(\thetab)$
equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data)  \mid \theta)
\end{equation}
where $c_\epsilon$ is a proportionality factor that depends on
$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
that is defined via a distance function $d$, i.e.\ $\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}$.

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$,
the second is the limit of $\epsilon \to 0$.  Whilst the probability
can be approximated with samples, this becomes computationally
infeasible if $\epsilon$ is too small. Hence, a large class of
inference methods work with $\epsilon >0$, which leads to the
approximate likelihood function $L_{d, \epsilon}(\thetab)$,
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
\end{equation}
Most often in likelihood-free inference, we are interested in the
posterior distribution of the parameters $\thetab$ given some observed
data. \textbf{HP comment: thinking about it, I think we're often more 
interested in a sample from the posterior, but this is maybe just semantics.} The approximate likelihood function in
\eqref{eq:approx_likelihood} implies the approximate posterior
$p_{d,\epsilon}(\thetab|\data)$,
\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

The approximation in \eqref{eq:approx_likelihood} is by no means the
only strategy to deal with the intractabilities of the likelihood
function in \eqref{eq:likelihood}. Other strategies include modelling
the (stochastic) relationship between $\thetab$ and $\yb$, and its
reverse, or framing likelihood-free inference as a ratio estimation
problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
  Thomas2016, Papamakarios2019, Chen2019, Thomas2020,
  Hermans2020}. However, both OMC and robust OMC, which we introduce
next, are based on the approximation in \eqref{eq:approx_likelihood}.

% likelihood of a specific parameter configuration $\thetab$ is equal to
% the probability that the simulator will produce outputs infinitively
% close to the observations $\data$, using this configuration. For
% continuous cases, the probability of perfectly replicating the
% observed data becomes zero. In this scenario a relaxation must be
% introduced by accepting simulated data that fall close to the data
% i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
% $\epsilon > 0$. The region can be defined as
% $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$ where
% $d(\cdot, \cdot)$ can represent any valid distance.

% This relaxation introduces the approximate likelihood,

% and the approximate posterior,