### 2022-02-07 latex_files/ch3/3-0-design-principles.tex

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI}. At a high
level, ROMC can be split into the training and the inference part. In
general terms, the training part covers all steps for estimating the
proposal regions and the inference part calculates the weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
Steps 2-10 (before the horizontal line) form the training part and
steps 12-17 the inference part, respectively.

\begin{algorithm}[!ht]
	\caption{ROMC. Requires the prior \( p(\thetab) \), the simulator
    \(M_r(\thetab)\), number of optimisation problems \(n_1\), number
    of samples per region \(n_2\), acceptance limit
    \(\epsilon\)}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{}
    \For{\(i \gets 1 \textrm{ to } n_1\)}
    \State \(\vb_i \sim p(\vb)\) \Comment{Draw nuisance variables}
    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \vb=\vb_i) \) \Comment{Define deterministic simulator}
      \State \( d_i(\thetab) = ||g(\thetab, \vb=\vb_i) - \yb_0||_2 \) \Comment{Define distance function}
      \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimisation problem}
      \If{\(d^*_i > \epsilon\)}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Estimate \(\mathcal{\hat{S}}_i\) and define \(q_i\) \Comment{Estimate proposal area}
      \State (Optional) Fit \(\Tilde{d}_i\) on \(\mathcal{\hat{S}}_i\) \Comment{Fit surrogate model}
      \\\hrulefill
      \For{\(j \gets 1 \textrm{ to } n_2\)}
      \State \(\thetab_{ij} \sim q_i\), compute \(w_{ij}\) as in Algorithm \ref{alg:sampling_GB} \Comment{Sample}
      \EndFor
    \EndFor
    \State \(E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
    \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalised posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions \(\mathcal{\hat{S}}_i\). The tasks are; (a) sampling
the nuisance variables \(\vb_i \sim p(\vb)\) for obtaining the
deterministic functions \(d_i(\thetab)\), (b) defining the optimisation
problems \(\min_{\thetab} \: d_i(\thetab)\), (c) obtaining
\(\thetab_i^*\), \(d_i^*\) (d) checking whether \(d_i^* \leq \epsilon\) and
(e) computing the proposal distribution \(q_i\).

If \(d_i(\thetab)\) is differentiable, using a gradient-based method
is advised for obtaining \(\thetab_i^*\) faster. In this case, the
gradients \(\nabla_{\thetab} d_i\) gradients are approximated
automatically with finite-differences. This approximation requires two
evaluations of \(d_i\) for \emph{each} parameter
\(\theta_d, d \in \{1, \ldots, D\}\)\footnote{\(D\) is the
  dimensionality of \(\thetab\), i.e.  \(\thetab \in \mathbb{R}^D\)},
which works in low-dimensional problems. If \(d_i(\thetab)\) is not
differentiable, Bayesian Optimisation can be used as an alternative
solution. In this scenario, the training part becomes slower due to
fitting of the surrogate model and the blind optimisation steps.

After obtaining the optimal points \(\thetab^*\), we estimate the
proposal regions. Algorithm~\ref{alg:region_construction} describes
the line search approach for finding the region's boundaries. An
important step for the proposal region estimation is deciding the axes
of the bounding box; the directions \(\vb_d, d = \{1, \ldots, D\}\) we
follow for reaching the boundaries. We approximate them as the
direction of the highest curvature of \(d_i\) at \(\thetab_i^*\). This
estimation is given by the eigenvalues of the Hessian matrix
\(\hessian_i\) of \(d_i\)\footnote{Either the real distance \(d_i\) or
  the Gaussian Process approximation \(\hat{d}_i\)} at
\(\thetab_i\). The Hessian matrix is approximated numerically. In case
where the distance function is the Euclidean, the Hessian matrix can
be also computed as \(\hessian_i = \jac_i^T\jac_i\), where \(\jac_i\)
is the Jacobian matrix of the summary statistics
\(\Phi(g(\thetab, \vb=\vb_i))\) at \(\thetab_i^*\). The approximation
through the Jacobian matrix has the computational advantage of using
only first-order derivatives.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
\( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\), (b) computing an
expectation \(E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
unnormalised posterior \(p_{d, \epsilon}(\thetab|\data)\). Sampling is
performed by getting \(n_2\) samples from each proposal distribution
\(q_i\). For each sample \(\thetab_{ij}\), the distance
function\footnote{As before, if a surrogate model \(\hat{d}\) is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing the expectation is easy after
weighted samples are obtained using the equation~\ref{eq:expectation},
so we do not discuss it separately. Evaluating the unnormalised
posterior requires the deterministic functions \(g_i\) and the prior
distribution \(p(\thetab)\)\footnote{There is no need for solving the
  optimisation problems and building the proposal regions.}. The
evaluation requires iterating over all \(g_i\) and evaluating the
distance from the observed data.

\begin{algorithm}[!ht]
	\caption{Approximation of \(\mathcal{\hat{S}}_i\) with a bounding box; 
  Requires: a model of distance \(d_i(\thetab)\), 
  an optimal point \(\thetab_i^*\), 
  a number of refinements \(K\), 
  a step size \(\eta\_\text{start}\), 
  maximum iterations \(M\) and 
  a curvature matrix \(\hessian_i\) (\(\jac_i^T\jac_i \) or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors \(\vb_{d}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,||\thetab ||)\)}
	\For{\(d \gets 1 \textrm{ to } D\)}
		\State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
		\State \(k \gets 0\)
		\State \(\eta \gets \eta\_\text{start}\) \Comment{Initialize \(\eta\)}
		\Repeat
          \State \(j \gets 0\)
        	\Repeat
            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}\) \Comment{Large step size \(\eta\).}
            \State \(j \gets j + 1\)
        	\Until{\(d(f_i(\Tilde{\thetab}), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}\)
        	\State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
        	\State \(k \gets k + 1\)
      \Until \(k = K\)
      \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d\) \Comment{Then, make the minimum step}
      \EndIf \label{algstep:box_constr_end}
    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
	\EndFor
	\State Fit a rectangular box around the region end points and define \(q_i\) as uniform distribution
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance \(d_i\), the prior distribution \(p(\thetab)\), the proposal distribution \(q_i\)}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \State \(\thetab_{ij} \sim q_i\)
          \If {\(d_i(\thetab_{ij}) > \epsilon\)}
            \State Go to 2 \Comment{Reject sample}
          \Else {}
            \State \(w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}\) \Comment{Compute weight}
            \State Store \((w_{ij}, \thetab_{ij})\) \Comment{Store weighted sample}
          \EndIf
    \end{algorithmic}
\end{algorithm}