### 2022-02-14 Appendix/1_3_notation.tex
reordered repo

In this section, we provide an overview of the symbols utilised in the
rest of the document. At this level, the quantities are introduced
quite informally; most of them will be defined formally in the next
chapters. We try to keep the notation as consistent as possible
throughout the document. The symbol $\R^N$, when used, describes that
a variable belongs to the $\text{N-dimensional}$ Euclidean space; $N$
does not represent a specific number. The bold formation ($\xb$)
indicates a vector, while ($x$) a scalar. Random variables are
represented with capital letters ($\Thetab$) while the samples with
lowercase letters ($\thetab$), i.e.\ $\thetab \sim \Thetab$.

\subsubsection*{Random Generator}
\label{sec:random-generator}
\begin{itemize}
\item $M_r(\thetab): \R^D \rightarrow \R$: The black-box data simulator
\end{itemize}

\subsubsection*{Parameters/Random Variables/Symbols}
\label{sec:variables}

\begin{itemize}
\item $D \in \mathbb{N}$, the dimensionality of the parameter-space
\item $\Thetab \in \R^D$, random variable representing the parameters of interest
\item $\data \in \R^N$, the vector with the observations
\item $\epsilon \in R$, the threshold setting the limit on the region
  around $\data$. When further notation is introduced
  regarding $\epsilon$\footnote{If
    $\epsilon_*, \: *:\{filter, region, cutoff\}$ values are not specified
    explicitly, they all share the common value of $\epsilon$}
  \begin{itemize}
  \item $\epsilon_{filter}$, threshold for discarding solutions
  \item $\epsilon_{region}$, threshold for building bounding box regions
  \item $\epsilon_{cutoff}$, threshold for the indicator function
  \end{itemize}

\item $\V \in \R^N$, random variable representing the randomness of
  the generator. It is also called nuisance variable, because we are
  not interested in inferring a posterior distribution on it.
\item $\vb_i \sim \V$, a specific sample drawn from $\V$
\item $\Y_{\thetab}$, random variable describing the simulator $M_r(\thetab)$. 
\item $\yb_i \sim \Y_\theta$, a sample drawn from $\Y_\theta$. It can
  be obtained by executing the simulator $\yb_i \sim M_r(\thetab)$
\end{itemize}

\subsubsection*{Sets}
\label{sec:sets}

\begin{itemize}
\item $\region(\data)$, the set of $\yb$ points close to the
  observations, i.e.\ $\yb := \{\yb: d(\yb, \data) \leq \epsilon \}$
\item $\regioni$, the set of points defined around $\yb_i$ i.e.\
  $\regioni = \region(\yb_i)$
\item $\mathcal{S}_i$, the set of parameters $\thetab$ that generate data close
  to the observations using the i-th deterministic generator, i.e.\
  $\{ \thetab: M_d(\thetab, \vb_i) \in \region(\data) \}$
\end{itemize}

\subsubsection*{Generic Functions}
\label{sec:generic-functions}

\begin{itemize}
\item $p(\cdot)$, any valid pdf
\item $p(\cdot | \cdot)$, any valid conditional distribution
\item $p(\thetab)$, the prior distribution on the parameters
\item $p(\vb)$, the prior distribution on the nuisance variables
\item $p(\thetab|\data)$, the posterior distribution
\item $p_{d,\epsilon}(\thetab|\data)$, the approximate posterior
  distribution  
\item $d(\mathbf{x}, \mathbf{y}): \R^{2N} \rightarrow \R$: any valid
  distance, the $L_2$ norm: $||\mathbf{x}-\mathbf{y}||_2$
\end{itemize}

\subsubsection*{Functions (Mappings)}
\label{sec:functions-mappings}

\begin{itemize}
\item $M_d(\thetab, \vb): \R^D \rightarrow \R$, the deterministic
  generator; all stochastic variables that are part of the data generation process are represented by the parameter $\vb$
\item $f_i(\thetab) = M_d(\thetab, \vb_i)$, deterministic generator associated with sample $\vb_i \sim p(\vb)$
\item $g_i(\thetab) = d(f_i(\thetab), \data)$, distance of the generated data $f_i(\thetab)$ from the observations
\item $T(\mathbf{x}): \mathbb{R}^{D_1} \rightarrow \mathbb{R}^{D_2}$
  where $D_1 > D_2$, the mapping that computes the summary statistic
\item $\indicator{\region(\data)}(\yb)$, the indicator function; returns 1 if $d(\yb, \data) \leq \epsilon$, else 0
\item $L(\thetab)$, the likelihood
\item $L_{d,\epsilon}(\thetab)$, the approximate likelihood
\end{itemize}