### 2022-02-14 article/latex_files/ch1/1_2_outline.tex

The remainder of the dissertation is organised as follows; in Chapter
2, we establish the mathematical formulation. Namely, we initially
describe the simulator-based models and provide some background
information on the fundamental algorithms proposed so far. Afterwards,
we provide the mathematical description of the ROMC approach
\autocite{Ikonomov2019}. Finally, we transform the mathematical
description to algorithms. In Chapter 3, we illustrate the
implementation part; we initially provide some information regarding
the Python package Engine for Likelihood-Free Inference (ELFI)
\autocite{1708.00707} and subsequently, we present the implementation
details of ROMC in this package. In general, the conceptual scheme
followed by the dissertation is

$$\text{Mathematical modelling} \rightarrow
\text{Algorithm} \rightarrow \text{Software}$$

In Chapter 4, we demonstrate the functionalities of the ROMC
implementation at some real-world examples; this chapter demonstrates
the accuracy of the ROMC method and our implementation's at
likelihood-free tasks. Finally, in Chapter 5, we conclude with some
thoughts on the work we have done and some future research ideas.