### 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (BO) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,0.310209700650407);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,0.366870797321359);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.4774021490772);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,0.847063787984617);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,0.950467021534753);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,1.63746647929829);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,0.63033836109258);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.01425623641141);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,1.15727348863503);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,1.43066001219801);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.26780484039086);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,1.57088642283557);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.26780484039086);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,1.90093404306951);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,1.86235774750728);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.72575321812577);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.01425623641141);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,1.84427294639855);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.78345203790796);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,3.1720295175941);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,2.35296949050992);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,3.68663899199996);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.71687198144523);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,2.49494901962233);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,2.27822359273028);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.42715377524836);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.27700840481905);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.4878209014166);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,1.99251141603071);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,1.75945430650476);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,1.79803060206703);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,1.78570494893043);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,1.72087801094253);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,1.75945430650479);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,1.51407154384225);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,1.00938102922818);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,1.1776112007662);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.672920686152122);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,1.00938102922817);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.841150857690152);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.504690514614091);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.168230171538028);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.607177584483422 0
0.607177584483422 3.5
};
\addlegendentry{$\mu = 0.607$}
\addplot [semithick, black, dashed]
table {%
0.427777567120938 1
0.758013118742592 1
};
\addlegendentry{$\sigma = 0.150$}
\end{axis}

\end{tikzpicture}