### 2022-02-15 article/latex_files/ch1/1-introduction.tex
final version

Simulator-based models are particularly captivating due to the
modelling freedom they provide. In essence, a simulator-based model can
describe any data generating mechanism that can be written as a finite
set of algorithmic steps. This modelling freedom comes at a cost;
performing the inference, i.e., sample or evaluate the posterior
distribution, is challenging.

Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
novel LFI approach for approximating the posterior
distribution. Afterwards, \citet{Forneron2016} provided a similar
method naming it `reverse sampler'. The central idea is turning the
stochastic data-generating mechanism into a set of
fully-parallelizable deterministic optimisation processes. In their
work, ~\citet{Ikonomov2019}, located some critical failure modes of
OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
the appropriate improvements.

In this paper, we present an extendible and parallelisable
implementation of ROMC at the \proglang{Python} package \pkg{Engine
  for Likelihood-Free inference (ELFI)}. We choose carefully the
designing principles for making the implementation extendible. As we
describe analytically in the next chapter, ROMC can be understood not
only as a specific algorithm, but mainly as a framework for obtaining
weighted samples from the posterior; it defines a sequence of
algorithmic steps, without enforcing a specific algorithm for solving
them. Therefore, a researcher may adopt ROMC's backbone approach, but
experiment with alternative methods to solve the tasks. We have
designed ROMC for facilitating such alterations. Finally, we have
tested the accuracy and the efficiency of our implementation on the
LFI examples supported by the \pkg{ELFI} package.

To the best of our knowledge, this is the first attempt to implement
the ROMC inference method to a generic LFI framework. Therefore, there
are not other implementation directly comparable to ours. For this
reason, we test the implementation against (i) an artificial example
with tractable likelihood and (ii) the second-order moving average
(MA2) example from the \pkg{ELFI} package. In the latter, we generate
data using known parameters and we compare our results with the
typical Rejection ABC approach.