### 2022-02-20 article/latex_files/ch4/4-4-extensibility.tex

ROMC describes a sequence of steps for approximating the posterior
distribution, without explicitly enforcing the methods that solve
these steps. Even though for each step a specific algorithm is
proposed by~\cite{Ikonomov2019}, in general ROMC is functional if a
practitioner thinks of an alternative way of approaching a specific
step. Considering this particularity, we designed the implementation
to support extensibility.

We have specified four specific points where a user may intervene with
their custom modules; (a) the gradient-based optimisation, (b) the
Bayesian Optimisation, (c) the proposal region construction and (d)
the surrogate model fitting. These are the four critical parts of the
ROMC procedure, whereas the rest of the code is the backbone of the
algorithm.

The four replaceable parts described above, are solved using the four
methods of the \linebreak \code{OptimisatioProblem} class; (a)
\code{solve_gradients(**kwargs)}, (b) \code{solve_bo(**kwargs)},
\linebreak (c) \code{build_region(**kwargs)}, (d)
\code{fit_local_surrogate(**kwargs)}. The user can create a custom
class that inherits the basic \code{OptimisatioProblem} and, then,
overwrite any of the four functions with custom ones. Suppose a user
wants to fit Neural Networks as local surrogate models
\(\tilde{d}_i\). The user should overwrite the
\code{fit_local_surrogate(**kwargs)} function with one that internally
trains a neural network. In the following snippet we illustrate that,
using the \code{neural_network.MLPRegressor} class of the
\pkg{scikit-learn} package.

\begin{Code}
---------------------------------- python ----------------------------------
class CustomOptim(OptimisationProblem):
    def __init__(self, **kwargs):
        super(CustomOptim, self).__init__(**kwargs)

    def fit_local_surrogate(self, **kwargs):
        nof_samples = 500
        objective = self.objective

        # helper function
        def local_surrogate(theta, model_scikit):
            assert theta.ndim == 1
            theta = np.expand_dims(theta, 0)
            return float(model_scikit.predict(theta))

        # create local surrogate model as a function of theta
        def create_local_surrogate(model):
            return partial(local_surrogate, model_scikit=model)

        local_surrogates = []
        for i in range(len(self.regions)):
            # prepare dataset
            x = self.regions[i].sample(nof_samples)
            y = np.array([objective(ii) for ii in x])

            # train Neural Network
            mlp = MLPRegressor(hidden_layer_sizes=(10,10), solver='adam')
            model = Pipeline([('linear', mlp)])
            model = model.fit(x, y)
            local_surrogates.append(create_local_surrogate(model))

        self.local_surrogates = local_surrogates
        self.state["local_surrogates"] = True
----------------------------------------------------------------------------
\end{Code}

In the same way, the user may replace each of the other three
functionalities. The only restriction that must be respected, concerns
the side-effects each method has at the \code{OptimisatioProblem}
class level. In the following four snippets we present the signature
of each method. We name the class-level variables that must be set
under the comment \code{\# side-effects}.

\begin{Code}
---------------------------------- python ----------------------------------
def solve_gradients(self, **kwargs):
  # custom solution procedure
  x = ...
  y = ...
  jac = ...
  hess_inv = ...

  # side-effects
  self.state["attempted"] = True
  if success:
    self.result = RomcOpimisationResult(x, y, jac, hess_inv)
    self.state["solved"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------
def solve_bo(self, **kwargs):
  # custom procedure
  x = ...
  y = ...
  custom_surrogate = ...

  # side-effects
  self.state["attempted"] = True
  if success:
    self.result = RomcOpimisationResult(x, y)
    self.surrogate = custom_surrogate
    self.state["solved"] = True
    self.state["has_fit_surrogate"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------
def build_region(self, **kwargs):
  # custom build_region method
  bounding_box: List[NDimBoudningBox] = ...
  success = True/False # whether region built correctly

  # side-effects
  self.eps_region = eps_region
  if success:
    # construct region
    self.regions = bounding_box
    self.state["region"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------
def fit_local_surrogate(self, **kwargs):
  # custom local surrogates
  custom_surrogates = ...
  success = True/False # whether local surrogates fit correctly

  # side-effects
  if success:
    self.local_surrogate = local_surrogates
    self.state["local_surrogates"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------    
\end{Code}

The two classes that may be needed for creating the custom routines
are \\ (a) \code{RomcOpimisationResult} and (b)
\code{NDimBoundingBox}. We present their signatures below.

\begin{Code}
---------------------------------- python ----------------------------------
class RomcOpimisationResult:
    def __init__(self, x_min, f_min, hess_appr):
        Parameters
        ----------
        x_min: np.ndarray (D,) or float, the minimum point
        f_min: float, distance at the minimum point
        hess_appr: np.ndarray (DxD), Hessian approximation at x_min
        """
----------------------------------------------------------------------------    
class NDimBoundingBox:
    def __init__(self, rotation, center, limits, eps_region):
        Parameters
        ----------
        rotation: np.array (D,D),  rotation matrix for the bounding box
        center: np.array (D,) center of the bounding box
        limits: np.ndarray, shape: (D,2), the limits of the bounding box
        eps_region: float, distance threshold 
----------------------------------------------------------------------------    
\end{Code}

% Let's say we have observed that the local area around $\theta_i^*$ is
% too complex to be represented by a simple quadratic model\footnote{as
%   in the current implementation}. Hence, the user selects a neural
% network as a good alternative. In the following snippet, we
% demonstrate how they could implement this enhancement without much
% effort; (a) they have to develop the neural network using the package
% of their choice (b) they must create a custom optimisation class which
% inherits the basic \code{OptimisationClass} and (c) they have to
% overwrite the \code{fit_local_surrogate} routine, with one that
% sets the neural network's prediction function as the
% \code{local_surrogate} attribute. The argument \code{**kwargs}
% may be used for passing all the important arguments, e.g.\ training
% epochs, gradient step etc. If, for example, they would like to set the
% size of the training set dynamically, we may replace \code{x =
%   self.regions[0].sample(30)} with \code{x =
%   self.regions[0].sample(kwargs["nof_examples"])}. Finally, they must
% pass the custom optimisation class, when calling the \code{ROMC}
% method.

% \begin{Code}
% ------------------------------ python snippet ------------------------------  
%   class NeuralNetwork:
%       def __init__(self, **kwargs):
%           # set the input arguments

%       def train(x, y):
%           # training code

%       def predict(x):
%           # prediction code

%   # Inherit the base optimisation class
%   class customOptim(elfi.methods.parameter_inference.OptimisationProblem):
%       def __init__(self, **kwargs):
%           super(customOptim, self).__init__(**kwargs)

%       # overwrite the function you want to replace
%       def fit_local_surrogate(**kwargs):
%           # init and train the NN
%           x = self.regions[0].sample(30) # 30 training points
%           y = [np.array([self.objective(ii) for ii in x])]
%           nn = NeuralNet()
%           nn.train(x,y)

%           # set the appropriate attribute
%           self.local_surrogate = nn.predict

%           # update the state
%           self.state["local_surrogate"] = True

%   # pass the custom inference method as argument
%   romc = elfi.ROMC(dist, bounds, custom_optim_class=customOptim)
% ----------------------------------------------------------------------------
% \end{Code}