### 2022-12-21 article/latex_files/ch4/4-1-training.tex

\begin{Code}
romc.solve_problems(n1, use_bo = False, optimizer_args = None)
\end{Code}

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimiser or the Bayesian optimisation (B0) scheme. The
three tasks are completed sequentially, as shown in
Figure~\ref{fig:romc_overview}. The definition of the optimisation
problems is performed by drawing \code{n1} integer numbers from a
discrete uniform distribution \(u_i \sim \mathcal{U}\{1,
2^{32}-1\}\). Each integer \(u_i\) is the seed used in \pkg{ELFI}'s
random simulator. The user may select the Bayesian Optimisation scheme
by setting the argument \code{use_bo = True} and pass custom arguments
to the optimizer through \code{optimize_args}.

\begin{Code}
romc.distance_hist(**kwargs)
\end{Code}

\noindent
This function helps the user decide which threshold \(\epsilon\) to
use, by plotting a histogram of the distances at the optimal point
\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in
case \code{use_bo=True}. The function accepts all keyword arguments
and forwards them to the underlying \code{pyplot.hist()} function of
the package \pkg{matplotlib}. In this way the user may customise some
properties of the histogram, such as the number of bins or the range
of values.

\begin{Code}
romc.estimate_regions(eps_filter, use_surrogate=None, fit_models=False)
\end{Code}

\noindent
This method estimates the proposal region around the optimal points,
following Algorithm~\ref{alg:region_construction}. By default, the
distance model that will be used follows the decision from the
previous step; if a gradient-based optimizer has been used, then the
real distance function \(d\) will be chosen. If BO, then then the
surrogate model \(\hat{d}\). In case, the user wants to enforce using
the original distance function \(d\) after BO, they may set
\code{use_surrogate=False}. Finally, the option \code{fit_models}
defines whether to fit local surrogate models \(\tilde{d}\) after
estimating the proposal regions.

\begin{Code}
romc.fit_posterior(args*)  # training in a single call
romc.visualize_region(i)   # acceptance region
romc.compute_eps(quantile) # estimates eps
----------------------------------------------------------------------------
\end{Code}

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \\ \code{estimate_regions} into a single
step. The function \code{visualize_region} can be used for plotting
the bounding box around the optimal point, when the parameter space is
up to 2D. The argument \code{i} is the index of the corresponding
optimisation problem. Finally, \code{compute_eps} returns the
appropriate distance value \(d_{i=\kappa}^*\) where
\(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number
of accepted solutions. It can be used to automate the selection of the
threshold \(\epsilon\), e.g.,
\code{eps=romc.compute_eps(quantile=0.9)}.

\subsubsection*{Example}

In the following snippet, we put together the routines described above
to perform the training part at our running example.

\begin{Code}
# Training (fitting) part
n1 = 500 # number of optimisation problems
seed = 21 # seed for solving the optimisation problems
use_bo = False # set to True for switching to Bayesian optimisation

# Training step-by-step
romc.solve_problems(n1 = n1, seed = seed, use_bo = use_bo)
romc.theta_hist(bins = 100) # plot hist to decide which eps to use

eps = .75 # threshold for the bounding box 
romc.estimate_regions(eps = eps) # build the bounding boxes

romc.visualize_region(i = 1) # for inspecting visually the bounding box

# Equivalent one-line command
# romc.fit_posterior(n1 = n1, eps = eps, use_bo = use_bo, seed = seed)
\end{Code}

\begin{figure}[ht]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}