### 2022-12-27 article/article.tex
revised until 2.2

\documentclass[article]{jss}
%% -- LaTeX packages and custom commands ---------------------------------------

\usepackage[usenames,dvipsnames]{xcolor}

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{orcidlink}
\usepackage{amsmath,amssymb,bbold, xfrac}

\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption} % for subfigure
\usepackage{float}

%% TikZ packages
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots}
\pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}

%% for algorithms
\usepackage{algorithm}
\usepackage{algpseudocode} % algorithmicx package

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{marginnote}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% Vasilis' custom commands
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\X}{\mathbf{x}}
\newcommand{\Z}{\mathbf{z}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\hess}{\mathbf{H}}
\newcommand{\Thetab}{\mathbf{\Theta}}

\newcommand{\vb}{\mathbf{v}}
\newcommand{\ub}{\mathbf{u}}
\newcommand{\yb}{\mathbf{y}}
\newcommand{\xb}{\mathbf{x}}

\newcommand{\jac}{\mathbf{J}}
\newcommand{\hessian}{\mathbf{H}}
\newcommand{\thetab}{\boldsymbol{\theta}}
\newcommand{\thetabij}{\thetab_{ij}}
\newcommand{\thetabi}{\thetab_i}

\newcommand{\simulator}{g}
\newcommand{\region}{B_{d,\epsilon}}
\newcommand{\indicator}[1]{\mathbb{1}_{#1}}
\newcommand{\regioni}{B_{d,\epsilon}^i}
\newcommand{\data}{\mathbf{y_0}}
\newcommand{\accregioni}{C^i_{\epsilon}}
\newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}

\newcommand{\danger}{\fontencoding{U}\fontfamily{futs}\selectfont\char 66\relax}
\newcommand{\TODO}[1]{\textcolor{red}{TODO #1}\marginpar{\textcolor{red}{\Large \danger}}}

\newcommand{\R}{\mathbb{R}}

\newcommand{\Ex}{\mathbb{E}}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

%% -- Article metainformation (author, title, ...) -----------------------------

\author{Vasilis Gkolemis~\orcidlink{0000-0002-2636-0245}\\ATHENA RC \And
  Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo}

\Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}

\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable ROMC implementation}

%% - \Abstract{} almost as usual
\Abstract{Performing inference in statistical models with an
  intractable likelihood is challenging, therefore, most
  likelihood-free inference (LFI) methods encounter accuracy and
  efficiency limitations. In this paper, we present the implementation
  of the LFI method Robust Optimisation Monte Carlo (ROMC) in the
  \proglang{Python} package \pkg{ELFI}. ROMC is a novel and efficient
  (highly-parallelisable) LFI framework that provides accurate
  weighted samples from the posterior. Our implementation can be used
  in two ways. First, a scientist may use it as an out-of-the-box LFI
  algorithm; we provide an easy-to-use API harmonised with the
  principles of \pkg{ELFI}, enabling effortless comparisons with the
  rest of the methods included in the package. Additionally, we have
  carefully split ROMC into isolated components for supporting
  extensibility. A researcher may experiment with novel method(s) for
  solving part(s) of ROMC without reimplementing everything from
  scratch. In both scenarios, the ROMC parts can run in a
  fully-parallelised manner, exploiting all CPU cores. We also provide
  helpful functionalities for (i) inspecting the inference process and
  (ii) evaluating the obtained samples. Finally, we test the
  robustness of our implementation on some typical LFI examples.}

%% - \Keywords{}
\Keywords{Bayesian inference, implicit models, likelihood-free, \proglang{Python}, \pkg{ELFI}}
\Plainkeywords{Bayesian inference, implicit models, likelihood-free, Python, ELFI}

%% - \Address{} of at least one author
\Address{
  Vasilis Gkolemis\\
  Information Management Systems Institute (IMSI)\\
  ATHENA Research and Innovation Center\\
  Athens, Greece\\
  E-mail: \email{vgkolemis@athenarc.gr}\\
  URL: \url{https://givasile.github.io}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------
\section{Introduction} \label{sec:intro}

Simulator-based models are particularly captivating due to the
provided modeling freedom. In essence, any data generating mechanism
that can be written as a finite set of algorithmic steps can be
programmed as a simulator-based model. In these cases, it is feasible
to generate samples using the simulator but it is infeasible to
evaluate the likelihood function. The intractability of the likelihood
makes the likelihood-free inference (LFI), i.e., the approximation of
the posterior distribution without using the likelihood function,
particularly challenging.

Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
novel LFI approach for approximating the posterior distribution. The
central idea is turning the stochastic data-generating mechanism into
a set of deterministic optimization processes. Afterwards,
\citet{Forneron2016} provided a similar method under the name `reverse
sampler'. In their work,~\citet{Ikonomov2019}, located some critical
limitations of OMC, so they proposed Robust OMC (ROMC), an alternative
version of OMC with the appropriate modifications.

In this paper, we present the implementation of ROMC at the
\proglang{Python} package \pkg{ELFI} (Engine for likelihood-free
inference)~\citet{1708.00707}. As we illustrate at
Section~\ref{sec:implementation}, we have carefully designed the
implementation to ensure extensibility. ROMC is a general framework
for obtaining weighted samples from the posterior, i.e., it defines a
sequence of algorithmic steps without enforcing a specific algorithm
for solving each step. Therefore, a researcher may use ROMC as the
backbone algorithm and develop novel methods to solve each separate
step.\footnote{For being a ready-to-use
algorithm,~\citet{Ikonomov2019} proposed a default method for each
step, but this choice is by no means restrictive.} We have designed
our software for facilitating such experimentation.

To the best of our knowledge, this is the first implementation of the
ROMC inference method to a generic LFI framework. We organize the
evaluation in three steps. First, for securing that our implementation
is accurate, we test it against an artificial example with a tractable
likelihood. The artificial example also serves as a step-by-step guide
for showcasing how to use the various functionalities of our
implementation. Second, we use the second-order moving average (MA2)
example from the \pkg{ELFI} package, using as ground truth the samples
obtained with Rejection ABC~\citet{lintusaari2017}, with a very large
number of trials. Finally, we present the execution times of ROMC,
measuring the speed-up obtained by using the parallel version of the
implementation.

% \clearpage
\section{Background}

We first give a short introduction to simulator-based models, we then
focus on OMC and its robust improvement, ROMC, and we, finally,
introduce \pkg{ELFI}, the underlying package that is used for the
implementation of ROMC.

\subsection{Simulator-based models and likelihood-free inference}

An implicit or simulator-based model is a parameterized stochastic
data generating mechanism. The key characteristic of simulator-based
models is that we can sample data points, but we cannot evaluate the
likelihood. Formally, a simulator-based model is a parameterized
family of probability density functions
\(\{ p(\yb|\thetab)\}_{\thetab}\) whose closed-form is either unknown
or computationally intractable. In these scenarios, we can only a
access the simulator \( m_r(\thetab) \), i.e., the black-box mechanism
(computer code) that generates samples \(\yb\) in a stochastic manner
from a set of parameters \(\thetab\). We denote the process of
obtaining samples from the simulator with
\( m_r(\thetab) \rightarrow \yb \). As shown by~\citet{Meeds2015}, it
is feasible to isolate the randomness of the simulator by introducing
a set of nuisance random variables denoted by \(\ub \sim
p(\ub)\). Therefore, for a specific tuple \((\thetab, \ub)\) the
simulator becomes a deterministic mapping \(g\), such that
\(\yb=\simulator(\thetab,\ub)\). In terms of computer code, the
stochasticity of a random process is governed by the global
seed. Although each software may handle the randomness in slightly
different ways\footnote{For example, at
  \pkg{Numpy}~\citet{harris2020array}, the pseudo-random number
  generation is based on a global state, whereas, in
  \pkg{JAX}~\citet{jax2018github} random functions consume a key that
  is passed as parameter. }, in all cases, freezing the initial seed
to a specific integer converts the simulation to a deterministic piece
of code.

Simulator-based models provide modeling freedom, including hidden
(unobserved) random variables or rule-based decisions. In fact, any
physical process that can be conceptualized as a computer program of
finite steps can be modeled as a simulator-based model. Hence,
implicit models are often used to model physical phenomena in the
natural sciences such as, e.g., genetics, epidemiology or
neuroscience. Further background on simulator-based models and example
applications can be found in the articles by~\citet{gutmann2016,
  lintusaari2017, sisson2018, cranmer2020}.

The modeling freedom of simulator-based models, however, comes at the
price of difficulties in inferring the parameters of
interest. Denoting the observed data as \(\data\), the main difficulty
lies at the intractability of the likelihood function
\(L(\thetab) = p(\data|\thetab)\). To better see the sources of the
intractability, and to address them, we go back to the basic
characterization of the likelihood as the (rescaled) probability of
generating data \(\yb\) that is similar to the observed data
\(\data\), using parameters \(\thetab\). Formally, the likelihood
\(L(\thetab)\) equals

\begin{equation}
  \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
\end{equation}
where \(c_\epsilon\) is a proportionality factor that depends on
\(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around
\(\data\) that is defined through a distance function \(d\), i.e., \
\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}\).

Equation~\ref{eq:likelihood} highlights the main source of
intractability; computing
\(\Pr(\simulator(\thetab,\ub) \in \region(\data)) | \thetab\) as the
fraction of samples that lie inside the \(\epsilon\) region around
\(\data\) is computationally infeasible in the limit of
\(\epsilon \to 0\). Hence, the constrainted is relaxed to
\(\epsilon > 0\), which leads to the approximate likelihood:

\begin{equation}
  \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
\end{equation}

and, in turn, to the approximate posterior:

\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data)
  \propto L_{d, \epsilon}(\thetab) p(\thetab)
  \propto L_{d, \epsilon}(\thetab) p(\thetab)
\end{equation}

The approximation in
Equations~\ref{eq:approx_likelihood},~\ref{eq:approx_posterior} is by
no means the only strategy to deal with the intractability of the
likelihood function in Equation~\ref{eq:likelihood}. Other strategies include
modeling the (stochastic) relationship between \(\thetab\) and
\(\yb\), and its reverse, or framing likelihood-free inference as a
ratio estimation problem, see for example \citet{blum2010, Wood2006,
  Papamakarios2016, Papamakarios2019, Chen2019, Thomas2020,
  Hermans2020}. However, both OMC and robust OMC, which we introduce
next, are based on the approximation in Equation~\ref{eq:approx_likelihood}.

\subsection{Optimization Monte Carlo (OMC)}

Our description of OMC~\citet{Meeds2015}
follows~\citet{Ikonomov2019}. We define the indicator function (boxcar
kernel) that equals one only if \(\xb\) lies in \(\region(\yb)\):
%
\begin{equation}
  \label{eq:indicator}
  \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise}
    \end{array} \right. \end{equation}
%

We, then, rewrite the approximate likelihood function
\(L_{d, \epsilon}(\thetab)\) of Equation~\ref{eq:approx_likelihood} as:

\begin{gather}
  \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) | \thetab) =
  \int_{\ub} \indicator{\region(\data)}(g(\thetab, \ub)) d \ub
\end{gather}

which can be approximated using samples from the simulator:

\begin{equation}
  \label{eq:samples_approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\simulator(\thetab, \ub_i))
 \quad \text{ where } \ub_i \sim p(\ub).
\end{equation}

In Equation~\ref{eq:samples_approx_likelihood_omc}, for each
\(\ub_i\), there is a region \(\accregioni\) in the parameter space
\(\thetab\) where the indicator function returns one, i.e.,
\(\accregioni = \{ \thetab: \simulator(\thetab, \ub_i) \in
\region(\data) \}\). Therefore, we can approximate the likelihood as:

\begin{equation}
  \label{eq:alt_view_theta}
  L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab)
\end{equation}

and the corresponding approximate posterior:

\begin{equation}
  \label{eq:approx_posterior_omc}
  p_{d,\epsilon}(\thetab|\data) \propto
  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
\end{equation}

As argued by \citet{Ikonomov2019}, these derivations provide a unique
perspective for likelihood-free inference by shifting the focus onto
the geometry of the acceptance regions \(\accregioni\). Indeed, the
task of approximating the likelihood and the posterior becomes a task
of characterising the sets \(\accregioni\). OMC by \citet{Meeds2015}
assumes that the distance \(d\) is the Euclidean distance
\(||\cdot||_2\) between summary statistics \(\Phi\) of the observed
and generated data, and that the \(\accregioni\) can be well
approximated by infinitesimally small ellipses. These assumptions lead
to an approximation of the posterior in terms of weighted samples
\(\thetab_i^*\) that achieve the smallest distance between observed
and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.,\
\begin{equation} \label{eq:omc_opt_prob}
\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2  , \quad \ub_i \sim p(\ub).
\end{equation}
The weighting for each \(\thetab_i^*\) is proportional to the prior
density at \(\thetab_i^*\) and inversely proportional to the determinant
of the Jacobian matrix of the summary statistics at \(\thetab_i^*\). For
further details on OMC we refer the reader to \citep{Meeds2015,
  Ikonomov2019}.

\subsection{Robust optimization Monte Carlo (ROMC)}

\citet{Ikonomov2019} showed that considering infinitesimally small
ellipses can lead to highly overconfident posteriors. We refer the
reader to their paper for the technical details and conditions for
this issue to occur. Intuitively, it happens because the weights in
OMC are only computed from information at \(\thetab_i^*\), and using
only local information can be misleading. For example if the curvature
of \(||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2\) at
\(\thetab_i^*\) is nearly flat, the curvature alone may wrongly
indicate that \(\accregioni\) is much larger than it actually is. In
our software package we implement the robust generalisation of OMC by
\citet{Ikonomov2019} that resolves this issue.

ROMC, firstly, approximates the acceptance regions \(\accregioni\),
and defines proposal distributions \(q_i(\thetab)\) on them. The
proposal distributions are used for generating posterior samples
\(\thetab_{ij} \sim q_i\). The samples are assigned (importance)
weights \(w_{ij}\) that compensate for using the proposal
distributions \(q_i(\thetab)\) and not the prior \(p(\thetab)\),
\begin{equation}
  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
  \label{eq:sampling}
\end{equation}
Given the weighted samples, any expectation
\(\E_{p(\thetab|\data)}[h(\thetab)]\) of some function \(h(\thetab)\), can be approximated as
\begin{equation} \label{eq:expectation}
  \E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}
\citet{Ikonomov2019} considered uniform distributions as proposal
distributions so that the main task is to approximate the acceptance
regions \(\accregioni\) and to represent them so that uniform sampling
is easy. The approximation of the acceptance regions contains two
compulsory and one optional step: (1) solving the optimisation
problems as in OMC, (2) constructing bounding boxes around
\(\accregioni\) and optionally, (3) refining the approximation via a
surrogate model of the distance.

\subsubsection*{Solving the deterministic optimisation problems}
For each set of nuisance variables \(\ub_i, i = \{1,2,\ldots,n_1 \}\),
we search for a point \(\thetab^*_i\) such that
\(d(\simulator(\thetab^*_i,\ub_i), \data) \le \epsilon\). In general,
\(d\) can be any valid distance, but for the rest of the paper we
consider \(d\) as the squared Euclidean distance. For notational
convenience, we denote \(d(\simulator(\thetab,\ub_i), \data)\) as
\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the
following optimisation problem:
\begin{subequations}
\begin{alignat}{2}
  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point
\(\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)\) only if
\(d_i(\thetab_i^*) \le \epsilon\). If \(d_i(\thetab)\) is
differentiable any gradient-based optimizer can be used
for~\ref{eq:optProb}. The gradients \(\nabla_{\thetab} d_i(\thetab)\)
can be either provided in closed form or approximated by finite
differences. In case \(d_i\) is not differentiable, Bayesian
Optimisation~\citep{Shahriari2016} provides an alternative
approach. In this scenario, apart from obtaining an optimal
\(\thetab_i^* \), a surrogate model \(\hat{d}_i(\thetab)\) of the
distance function \(d_i(\thetab)\) is also automatically obtained;
\(\hat{d}_i\) can then substitute the actual distance function in
downstream steps of the algorithms, with possible computational gains
especially if evaluating the actual distance \(d_i(\thetab)\) is
expensive.

\subsubsection*{Estimating the acceptance regions}
The acceptance region \(\accregioni\) is approximated by a bounding
box \(\accregionihat\). Ideally, we want the bounding box to be as
tight as possible to \(\accregioni\) to ensure high acceptance rate in
the importance sampling, but big enough for not discarding valid
parts. The bounding boxes are built in two steps. First, we define
their axes \(\mathbf{v}_m\), \(m = \{1, \ldots, D\}\) based on the
(estimated) curvature of the distance at \(\thetab_i^*\). Second, we
determine the size of the box via a one-dimensional line-search method
along each axis, see Algorithm~\ref{alg:region_construction} for the
details. After the bounding boxes construction, a uniform distribution
\(q_i\) is defined on each bounding box, and is used as the proposal
region for importance sampling.

\subsubsection*{Refining the estimate via a local surrogate model (optional)}
When computing the weight \(w_{ij}\) in \eqref{eq:sampling}, we need
to check whether the samples \(\thetab_{ij} \sim q_i\) lie inside the
acceptance region \(\accregioni\). This can be considered to be a
safety-mechanism that corrects for any inaccuracies in the
construction of \(\accregionihat\) above. However, this check involves
evaluting the distance function \(d_i(\thetab_{ij})\), which can be
expensive if the model is complex. \citet{Ikonomov2019} thus proposed
to fit a surrogate model \(\tilde{d}_i(\thetab)\) of the distance
function \(d_i(\thetab)\), on data points that lie inside
\(\accregionihat\). They used a simple quadratic model whilst other
regression models are, in principle, possible too. The advantage of
using a quadratic model is that it has ellipsoidal isocontours, which
thus naturally allowed \citet{Ikonomov2019} to replace the bounding
box approximation of \(\accregioni\) with a tigher-fitting ellipsoidal
approximation.\footnote{The difference to the infinitesimal
  ellipsoidal model in OMC is the estimation procedure: OMC uses
  information at \(\thetab_i^*\) whilst, here, information in
  \(\accregionihat\) is used, which results in a more stable fit.}

The training data for the quadratic model is obtained by sampling
\(\thetab_{ij} \sim q_i\) and accessing the distances
\(d_i(\thetab_{ij})\). The generation of the training data adds an
extra computational cost, but leads to a significant speed-up when
evaluating the weights \(w_{ij}\). Moreover, the extra cost is largely
eliminated if Bayesian Optimisation with a Gaussian process (GP)
surrogate model \(\hat{d}_i(\thetab)\) was used to obtain
\(\thetab_i^*\) in the first step. In this case, we can use
\(\hat{d}_i(\thetab)\) instead of \(d_i(\thetab)\) to generate the
training data. This essentially replaces the global GP model with a
simpler local quadratic model which is typically more robust.

\subsection[Engine for likelihood-free inference (ELFI)]{\pkg{Engine for likelihood-free inference (ELFI)}}
\label{subsec:ELFI}
\input{latex_files/ch2/2-4-ELFI.tex}

%% -- Design Principles -----------------------------------------------------
\section{Implementation}
\label{sec:implementation}
This section is split in two parts. We first express ROMC as an
algorithm and then we present the general implementation principles we
follow.

\subsection{Algorithmic view of ROMC}
\input{latex_files/ch3/3-0-design-principles.tex}

\subsection{General implementation principles}
\label{subsec:general_design}
\input{latex_files/ch3/3-1-implementation-principles.tex}

%% -- ROMC implementation --
\section{Implemented functionalities}
\input{latex_files/ch4/4-0-intro.tex}

\subsection{Training part}
\label{subsec:training}
\input{latex_files/ch4/4-1-training.tex}

\subsection{Inference part}
\label{subsec:inference}
\input{latex_files/ch4/4-2-inference.tex}

\subsection{Evaluation part}
\label{subsec:evaluation}
\input{latex_files/ch4/4-3-evaluation.tex}

\subsection{Extensibility} % better title
\label{subsec:extensibility}
\input{latex_files/ch4/4-4-extensibility.tex}

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}
\input{latex_files/ch5/5-ma2.tex}

%% -- Summary/conclusions/discussion -------------------------------------------
\section{Summary and discussion} \label{sec:summary}
\input{latex_files/ch6/6-summary.tex}

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

The results in this paper were obtained using \proglang{Python}~3.7.9
with the \pkg{ELFI}~0.8.3 package. The experiments have been executed
in a single machine with an Intel® Core™ i7-8750H Processor an with
Ubuntu 20.04 lts operating system.

\section*{Acknowledgments}

HP was funded by European Research Council grant 742158 (SCARABEE,
Scalable inference algorithms for Bayesian evolutionary epidemiology).

\clearpage
\bibliography{refs}

%% -----------------------------------------------------------------------------

\end{document}